### 2021-11-04 paper/main.tex
add paper to repo

\documentclass{article}
\usepackage{caption}
\usepackage{wrapfig}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage{comment}
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{xurl}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
%

\usepackage{microtype}      % microtypography
\usepackage{todonotes}
\usepackage{longtable}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{amsmath}
\usepackage{utils/kky}
\usepackage{subcaption}
\usepackage{afterpage}
\usepackage{float}
\usepackage{ bbold }
\title{FastSRM: A fast, memory efficient and identifiable implementation of the
  shared response model}
\date{\today} % clear date

\begin{document}
\maketitle

\begin{abstract}
  The shared response model (SRM) provides a simple but effective framework to analyse
  fMRI data of subjects exposed to naturalistic stimuli.
  %
  However when the number
of subjects or runs is large, fitting the model requires a large amount of memory
and computational power, which limits its use in practice.
%
Furthermore, SRM is
not identifiable which makes the shared response difficult to interpret.
%

In this work, we implement an identifiable version of SRM and show on real data
that it improves the stability of the recovered shared response.
%
We then introduce the FastSRM that relies on a dimension reduction step that we
prove to be optimal in the sense that working with reduced data does not induce any change in the
algorithm trajectory.
%
We show experimentally using synthetic and real fMRI data
that FastSRM is considerably faster and more memory efficient
than competitive implementations.
%

The experiments performed in the paper are \emph{fully} reproducible: our
code available at \url{https://github.com/hugorichard/FastSRM} downloads the
data, runs the experiments and plot the figures.
%

\end{abstract}

\section{Introduction}
What makes the analysis of functional magnetic resonance imaging (fMRI) data
particularly difficult is that the number of features (voxels) is usually much
larger than the number of available samples (timeframes).
%
In such cases, dimension reduction methods are precious tools that allow to
reduce the computational burden of subsequent data analysis.
%
The shared response
model~\cite{chen2015reduced} provides a principled way to perform dimension
reduction of brain imaging data from subjects performing the same task and is
therefore a natural basis for common source
identification~\cite{richard2021model}~\cite{richard2020modeling},
classification~\cite{turek2018capturing}~\cite{chen2017shared}\cite{zhang2016searchlight}
or transfer learning~\cite{zhang2018transfer}.
%

However SRM has initially been designed to work within regions of interest using
only few subjects.
%
When using full brain data, computational costs become
high.
%
In addition, memory requirements are difficult to meet since the full dataset
needs to hold in memory.
%
Another issue of SRM is its non-identifiability (see the proof
in~\cite{richard2020modeling} Appendix D) which makes it challenging to
interpret the extracted shared components as they are defined up to a rotation.
%

In this work, we introduce FastSRM, an identifiable model that uses an optimal
atlas based decomposition to speed up the computations with provably no loss of
performance.
%
FastSRM gains identifiability by imposing that the covariance of the shared
components is diagonal (see ~\cite{richard2020modeling} Appendix D for a proof).
%
The gains in speed and memory usage are based on the intuition that a compressed
representation of the input should be good enough to find a suitable estimate of
the shared components as they live in reduced space.
%
It turns out that there
exists an optimal atlas that reduces the data in a such a way that the obtained
shared components (and other parameters) are the same as if they were computed
on full data.
%
We show that FastSRM brings several practical benefits: on real data, it produces more stable
estimates of the shared components than its non-identifiable counter-part and 
on synthetic as well as on real data, it is much faster and more memory
efficient than other implementations that do not make use of the optimal atlas.
%

Our python implementation is available at
\url{https://github.com/hugorichard/FastSRM}.
%
All our experiments are fully reproducible.
%
Scripts to download the data, run the experiments and plot the figures are
included.
%
A continuous integration pipeline runs the tests
automatically when any change to the core algorithm is made.
%

\section{Background: the shared response model (SRM)}
\label{sec:srm:review}
The shared response model~\cite{chen2015reduced} is a multi-view latent factor
model.
%
The data $\xb_1 \dots \xb_m$ are modeled as random vectors following the model:
\begin{align}
 &\xb_i = A_i \sbb + \nb_i \\
  &A_i^{\top}A_i = I_p
  \label{eq:model:srm}
\end{align}
where $\xb_i \in \RR^v$ is the data of view $i$, $A_i \in \RR^{p \times v}$ is the
mixing matrix of view $i$, $\nb_i$ is the noise of view $i$ and $\sbb \in \RR^p$ are the
shared components referred to as the \emph{shared response} in fMRI
applications.
%

The mixing matrices
$A_i$ are assumed to be orthogonal so that $A_i^{\top}A_i = I_p$.
%
However, in
general the matrix $A_i A_i^{\top}$ is different from identity.
%
The noise
$\nb_i$ is assumed to be Gaussian with covariance $\Sigma_i$ and independent
across views.
%
We assume the number of features $v$ to be much larger than the
number of components $p$: $v \gg p$.
%

The conceptual figure~\ref{fig:srm:conceptual_figure} illustrates an 
application of the shared response model to fMRI data.
%
The mixing
matrices are spatial topographies specific to each subjects while the shared
components give the common timecourses.
%
In~\cite{chen2015reduced, anderson2016enabling}, two versions of the shared
response model are introduced which we now present.
%

\begin{figure}
  \centering
  \includegraphics[scale=0.3]{figures/srm/conceptual_figure31.png}
  \caption{\textbf{Shared response model}: The raw fMRI data are modeled as a
    weighted combination of subject-specific spatial components with additive
    noise.
    %
    The weights are shared between subjects and constitute the shared response
    to the stimuli.
    %
  }
  \label{fig:srm:conceptual_figure}
\end{figure}

\subsection{Deterministic shared response model}
\label{sec:deterministicsrm}
Let us consider $n$ observations of $\xb_i$ and $\sbb$ that we stack into
matrices $X_i \in \RR^{v, n}$ and $S \in \RR^{p, n}$.
%

The deterministic shared response model sees both the mixing matrices $A_i$ and
the $n$ observations of the shared response $S$ as parameters to be
estimated.
%
The noise variance is fixed to a multiple of identity: $\forall i,
\Sigma_i=\sigma^2 I_v$ where $\sigma$ is an hyper-parameter to choose.
%

The model is optimized by maximizing the log-likelihood.
%

The likelihood is given by: $p(\xb) = \prod_i \Ncal(\xb_i; A_i \sbb, \sigma^2 I)$ and
therefore the empirical expected negative log-likelihood is given up to a constant independent of
$A_i$ and $S$ by:
\begin{align}
  \loss = \frac1{n} \sum_i \|A_i S - X_i \|^2 = \frac1{n} \big(\| S \|^2 -2 \dotp{ A_i S}{ X_i } + \| X_i \|^2 \big)
  \label{eq:detsrmloss}
\end{align}
$\loss$ is optimized by performing alternate minimization on $(A_1 \dots A_m)$
and $S$.
%
Note that the hyper-parameter $\sigma$ does not have an influence on
the loss and can therefore be safely ignored.
%

The gradient with respect to $S$ is given by $\sum_i A_i^{\top}(A_i S -
X_i) = \sum_i (S -
A_i^{\top} X_i)$
yielding the closed form updates:
\begin{equation}
  S \leftarrow  \frac1m \sum_i (A_i^{\top} X_i)
  \label{eq:srm:supdate}
\end{equation}

From \eqref{eq:detsrmloss}, minimizing $\loss$ with respect to $A_i$ is
equivalent to maximizing $\dotp{ A_i}{ X_i S^{\top} }$ and therefore we
have:
\begin{equation}
  A_i \leftarrow  \Pcal(\frac1{n}X_i S^{\top})
  \label{eq:detsrm:Aiupdate}
\end{equation}
where $\Pcal$ is the projection on the Stiefel manifold: $\Pcal(M) = M
(M^{\top}M)^{-\frac12}$.
%

The complexity of Deterministic SRM is in $\bigO(\mathrm{n_{iter}} mpvn)$ where
$n$ is the number of samples and $\mathrm{n_{iter}}$ the number of iterations.
%
We monitor the convergence by looking at the $\ell_{\infty}$ norm of the
gradient.
%
Note that we can monitor the gradient without any increase in complexity.
%
Indeed, after the updates with respect to each mixing matrix have been
carried out, only the gradient with respect to $S$ remains: $\sum_i
(S - A_i^{\top}\xb_i)$.
%
The algorithm is stopped when the
gradient falls below a chosen tolerance.
%

\subsection{Probabilistic SRM}
\label{sec:probabilisticsrm}
In Probabilistic SRM , $\Sigma_i=\sigma_i^2 I_v$ and the shared
components are assumed to be Gaussian $\sbb \sim \Ncal(0, \Sigma_s)$.
%

In~\cite{chen2015reduced} and~\cite{anderson2016enabling}, $\Sigma_s$ is only assumed to be definite positive.
%
As already highlighted in introduction, this causes the model to be
unidentifiable (see~\cite{richard2020modeling} Appendix D for a proof). Enforcing a diagonal $\Sigma_s$ ensures identifiability (provided the diagonal values are different).
%
So we assume here that $\Sigma_s$ is diagonal (and refer the interested reader
to~\cite{chen2015reduced} and~\cite{anderson2016enabling} for the original formulation of Probabilistic SRM
without the diagonal constraint)
%

The model is optimized via the expectation maximization algorithm.
%
Denoting $\VV[\sbb | \xb] = (\sum_i \frac1{\sigma_i^2} I +
\Sigma_s^{-1})^{-1}$ and $\EE[\sbb | \xb] = \VV[\sbb | \xb] \sum_i \frac1{\sigma_i^2}
A_i^{\top}\xb_i$, we have
\begin{align}
  p(\xb, \sbb) &= \prod_i \frac{\exp(-\frac{\|\xb_i - A_i \sbb \|^2}{2 \sigma_i^2})}{(2 \pi \sigma_i^{2v})^{\frac12}} \frac{\exp(-\frac12 \dotp{ \sbb }{ \Sigma_s^{-1} \sbb } )}{(2 \pi | \Sigma_s|)^{\frac12}} \\
               &= c_1 \exp(-\frac12 \left( \sum_i \frac1{\sigma_i^2}\|\xb_i\|^2 - 2  \dotp{ \sum_i \frac1{\sigma_i^2} A_i^{\top}\xb_i}{ \sbb } \right. \\& \left.+ \sum_i \frac1{\sigma_i^2} \| \sbb \|^2 + \dotp{ \sbb}{ \Sigma_s^{-1} \sbb }  \right)) \\
               &= c_2(\xb) \exp(-\frac12 \left( \dotp{  \sbb - \EE[\sbb }{ \xb], \VV[\sbb | \xb]^{-1} ( \sbb - \EE[\sbb | \xb])  } \right)) \label{probsrmcompleted}
\end{align}
where
\begin{align}
  c_1 = \frac1{(2 \pi \sigma_i^{2v})^{\frac12}}\frac1{(2 \pi |
  \Sigma_s|)^{\frac12}}
\end{align}
and
\begin{align}
  c_2(\xb) = c_1 \exp(-\frac12( \sum_i
  \frac1{\sigma_i^2}\|\xb_i\|^2 - \dotp{  \EE[\sbb }{ \xb], \VV[\sbb | \xb]^{-1}
  \EE[\sbb | \xb] }))
\end{align}
are independent of $\sbb$.
%
Therefore,
\begin{align}
  \sbb| \xb \sim \Ncal(\EE[\sbb | \xb], \VV[\sbb, \xb])
\end{align}

The negative expected completed log-likelihood is given by
\begin{align}
	\loss = \sum_i \frac12 v \log(\sigma_i^2) + \frac1{2 \sigma_i^2} \EE[\| \xb_i - A_i \sbb \|^2]
\end{align}
updates are therefore given by:
\begin{align}
&\sigma_i^2 \leftarrow \frac1{v} (\EE[\| \xb_i - A_i \EE[\sbb|\xb]\|^2] + \| \diag(\VV[\sbb | \xb]) \|^2) \\
  &A_i \leftarrow \Pcal(\EE[\xb_i \EE[\sbb|\xb]^{\top}]) \label{eq:psrm:Aiupdate} \\
  & \Sigma_s \leftarrow \VV[\sbb | \xb] + \EE[\EE[\sbb | \xb] \EE[\sbb | \xb]^{\top}]
\end{align}

It is useful to access the log-likelihood to check the implementation of the
algorithm and monitor the convergence.
%
From equation~\eqref{probsrmcompleted},
the likelihood is given by:
\begin{align}
  p(\xb) &= c_2(\xb) \int_{\sbb} \exp(-\frac12 \left( \dotp{  \sbb - \EE[\sbb }{ \xb], \VV[\sbb | \xb]^{-1} ( \sbb - \EE[\sbb | \xb])  } \right)) d\sbb \\
         &= c_2(\xb) (2 \pi |\VV[\sbb | \xb]|)^{\frac12}
\end{align}
replacing $c_2(\xb)$ by its expression and taking the log, the expected negative
log-likelihood is (up to constants) given by:
\begin{align}
  \EE[-\log(&p(\xb))] = \sum_i \frac{v}{2} \log(\sigma_i^2) + \frac12 \log(|\Sigma_s|) - \frac12 \log(|\VV[\sbb | \xb]|) \nonumber \\ &+ \sum_i
  \frac12 \frac1{\sigma_i^2} \EE[\|\xb_i\|^2] - \frac12 \EE[\dotp{  \EE[\sbb }{ \xb], \VV[\sbb | \xb]^{-1} \EE[\sbb | \xb] }]
\end{align}

The complexity of Probabilistic SRM is $\bigO(\mathrm{n_{iter}} mpvn)$, the same as in
Deterministic SRM.
%
We can monitor the convergence by looking at the log-likelihood decrease at each iteration
and stop the algorithm when the magnitude of the decrease is below some
tolerance.
%
The storage requirements of Deterministic or Probabilistic SRM are in
$\bigO(mvn)$ which simply means that the dataset needs to hold in memory.
%

\section{The FastSRM algorithm}

\subsection{Reducing the computational burden by use of atlases}
SRM algorithms use different set of parameters $\theta$ to represent the data.
%
In deterministic SRM $\theta = (A_i)_{i=1}^m, S$ where $(A_i)_{i=1}^m$ are the
mixing matrices and $S$ are the $n$ observations of the shared response $S$ while in probabilistic SRM $\theta
= (A_i)_{i=1}^m, \Sigma_s, (\sigma_i)_{i=1}^m$ where $(A_i)_{i=1}^m$ are the
mixing matrices, $(\sigma_i)_{i=1}^m$ the noise levels and $\Sigma_s$ the
components variance.
%

In fMRI, the classical approach used to reduce the data is to apply an atlas.
%
A deterministic atlas such as~\cite{bellec2010multi} is a parcellation of the
brain into $r$ regions.
%
Reducing an image using a deterministic atlas corresponds to
averaging the signal within each region of the atlas.
%
A probabilistic atlases such
as~\cite{dadi_fine-grain_2020} describes each region as a set of weights across
the full brain.
%
Therefore, the image reduction can be done with a matrix product.
%

In FastSRM we consider a set of view specific atlases $U_i \in \RR^{v \times r}$ such that
$U_i^{\top}U_i = I_r$ where $r$ is the number of regions in the atlas.
%
Data are reduced using $\zb_i = U_i^{\top} \xb_i$ and an SRM algorithm is applied
on data $\zb_i$ yielding parameters $\theta'$.
%
The figure~\ref{fig:srm:conceptual} illustrates this process.
%

Note that the parameters obtained with FastSRM $\theta'$ are different
from the parameters obtained with the corresponding SRM algorithm $\theta$ (the
unmixing matrices in $\theta'$ do not even have the same shape as the unmixing
matrices in $\theta$).
%
However, as we will see in the next section, there exists atlases such as the
correspondence between $\theta$ and
$\theta'$ can be made explicit.
%

\begin{figure}
  \centering
  \includegraphics[scale=0.34]{figures/srm/conceptual_figure2.pdf}
  \caption{\textbf{FastSRM algorithm} In step 1, data $\xb_i$ are projected onto an
    atlas $U_i$ that may depend on the subject $i$ (top).
    %
    In step 2 a SRM algorithm is applied on reduced data to compute the shared
    response.
    %
  }
  \label{fig:srm:conceptual}
\end{figure}

From a computational stand point, the dimension reduction provides
a large reduction in memory usage.
%
Indeed as the original data are seen only
once, it is no longer necessary to keep the full dataset in memory (we can load
data $X_i$ one after the other and similarly for the atlases $U_i$).
%
Therefore
the memory consumption is only in $\bigO(vn)$ (where $v$ is the number of voxels
and $n$ is the number of samples) which is lower than SRM by a factor of $m$,
the number of subjects.
%
The number of subjects is typically between $10$ and
$1000$.
%
This yields a practical benefit: on fMRI datasets with many subjects, one no
longer needs a large cluster to run the shared response model but only a modern
laptop.
%
Additionally, low memory consumption reduces the
risk of thrashing~\cite{denning1968thrashing}, a phenomenon that causes large
increase in computation time when the memory used is close to the total available
memory in the hardware.
%

After preprocessing, the reduced representation $\zb_i$ is used instead of the
original data $\xb_i$ yielding a time complexity of $\bigO(
\mathrm{T_{preprocessing}} + \mathrm{n_{iter}} mpnr)$.
%
Let us highlight that an experiment is often run
multiple times such as when cross validated results are needed.
%
In these cases,
the pre-processing is performed only once and the apparent complexity becomes
$\bigO(\mathrm{n_{iter}} mpnr)$ which is faster than SRM by
a factor of $\frac{v}{r}$.
%
The number of regions in large atlases is about $r=1000$ and in full brain data,
the number of voxels is about $300~000$ so that $\frac{v}{r}$ is typically about
$1000$.
%
It remains to show how to draw a correspondence between FastSRM and SRM which is
addressed in the following section.
%

\subsection{An optimal atlas}
\label{sec:optimal_atlas}
In principle, FastSRM can be used with any atlas.
%
However, in general, working with reduced data induces a loss of information
that can be minimized if the atlas is carefully chosen.
%
In any case, there is little hope to recover the parameters that would have been
obtained from SRM from the parameters of FastSRM.
%
Yet, we show that there exists an optimal atlas in the
sense that SRM and FastSRM yield the same results.
%

Let us consider $\xb_i = U_{\xb_i} \zb_i$ a PCA of $\xb_i$ with the maximum
number of components.
%
As the number of
samples $n$ is lower than the number of features, $U_{\xb_i} \in \RR^{v \times
  n}$ and $\zb_i \in \RR^{n}$.
%
We also have $U_{\xb_i}^{\top}U_{\xb_i} = I$.
%
Therefore $U_{\xb_i}$ constitutes a possible choice of subject specific atlas.
%
As the next property shows, $U_{\xb_i}$ is an optimal atlas for deterministic
FastSRM.
%

\begin{proposition}[Optimal atlas for deterministic FastSRM]
  Let $(A_i)_i, S$ be the solution obtained by deterministic SRM on data
  $(X_i)_i$ and $(A_i')_i, S'$ the solution obtained by deterministic FastSRM on
  data $(X_i)_i$ using atlases $(U_{X_i})_i$ where $X_i = U_{\xb_i} Z_i$ is a PCA
  of $X_i$.
  %
  Then $A_i = U_{X_i}A_i'$ and $S = S'$.
  %

\label{prop:optimaldetsrm}
\end{proposition}
\begin{proof}
Updates of the mixing matrices $A_i$ in deterministic SRM
equation~\eqref{eq:detsrm:Aiupdate} can be written:
\begin{align}
  A_i &\leftarrow \Pcal(\frac1{n}X_i S^{\top})= U_{X_i}\Pcal(\frac1{n}Z_i S^{\top})
  \label{eq:fastdetsrm:Aiupdate}
\end{align}
where $\Pcal$ is the projection on the Stiefel manifold: $\Pcal(M) = M
(M^{\top}M)^{-\frac12}$.
%

Therefore we can look for $A_i$ as $A_i = U_{X_i} \tilde{A}_i$.
%
We can show that $\tilde{A}_i$ is orthogonal.
%
Indeed,
\begin{align}
  &A_i^{\top} A_i = I_p \\
  & \implies \tilde{A}_i^{\top}U_{X_i}^{\top} U_{X_i} \tilde{A}_i = I_p \\
  & \implies \tilde{A}_i^{\top} \tilde{A}_i = I_p
\end{align}

Then, we use the fact that
\begin{align}
  \|X_i - A_i S \|^2 = \| U_{X_i}Z_i - U_{X_i}\tilde{A}_i S\|^2 = \| Z_i - \tilde{A}_i S \|^2
  \label{eq:equality:xy}
\end{align}
so that $A_i' = \tilde{A}_i$.
%

Therefore, the solution of deterministic SRM on data $(\zb_i)_{i=1}^m$ and
$(\xb_i)_{i=1}^m$ are linked by the change of parameters $A_i = U_{\xb_i}A_i'$ and
$S = S'$.
%
This concludes the proof.
%

\end{proof}

In the case of probabilistic SRM we can obtain very similar results.
%
However the
algorithm applied on reduced data need to be slightly modified.
%

We call probSRM($\psi$) the probabilistic SRM algorithm modified such that
updates
\begin{align}
\sigma_i^2 \leftarrow \frac1{v} (\EE[\| \xb_i - A_i \EE[\sbb|\xb]\|^2] + \| \diag(\VV[\sbb | \xb]) \|^2)
\end{align}
are replaced by updates
\begin{align}
  \sigma_i^2 \leftarrow \frac1{\psi} (\EE[\| \xb_i - A_i \EE[\sbb|\xb]\|^2] + \| \diag(\VV[\sbb | \xb]) \|^2)
\end{align}

We have the following result:
\begin{proposition}[Optimal atlas for probabilistic FastSRM]
  Let $(A_i)_i, \sigma_i, \Sigma_s$ be the solution obtained by probabilistic SRM on data
  $\xb_i$ and $(A_i')_i, \sigma_i', \Sigma_s'$ the solution obtained by ProbSRM($v$) on
  data $\zb_i = U_{\xb_i}^{\top} \xb_i$.
  %
  Then $A_i = U_{\xb_i}A_i'$, $\sigma_i =
  \sigma_i'$ and $\Sigma_s = \Sigma_s'$.
  %

  \label{prop:optimalprobsrm}
\end{proposition}
\begin{proof}
  Updates of the mixing matrices $A_i$ in probabilistic SRM
  equation~\eqref{eq:psrm:Aiupdate} can be written:
  \begin{align}
    A_i &\leftarrow U_{\xb_i}\Pcal(\EE[\zb_i \EE[\sbb| \xb_i]^T])
    \label{eq:fastprobsrm:Aiupdate}
  \end{align}
  so we can look for $A_i$ as $A_i = U_{\xb_i} \tilde{A}_i$ and, as in the
  deterministic case, $\tilde{A}_i$ is orthogonal.
  %
  Therefore equality~\eqref{eq:equality:xy} holds.
  %

  Then we consider the expected negative log-likelihood of probabilistic srm:
  \begin{align}
    \loss &= \sum_i \frac12 v\log(\sigma_i^2) + \frac12 \log(|\Sigma_s|) + \EE[\int_{\sbb} \sum_i \frac1{2 \sigma_i^{2}} \|\xb_i - A_i \sbb \|^2 \nonumber \\& \enspace \enspace \enspace \enspace+ \frac12 \dotp{ \sbb }{ \Sigma_s^{-1} \sbb }  d\sbb] \\
          &= \sum_i \frac12 v \log(\sigma_i^2) + \frac12 \log(|\Sigma_s|) + \EE[\int_{\sbb} \sum_i \frac1{2 \sigma_i^{2}}\|\zb_i - \tilde{A}_i \sbb \|^2 \nonumber \\& \enspace \enspace \enspace \enspace+ \frac12 \dotp{ \sbb }{ \Sigma_s^{-1} \sbb }  d\sbb]
  \end{align}
  where we use equality~\eqref{eq:equality:xy}.
  %

  Optimizing the log-likelihood via expectation maximization yields the exact
  same updates as probabilistic SRM on data $\zb_i$
  except that updates
  \begin{align}
    \sigma_i^2 \leftarrow \frac1{t} (\EE[\| \zb_i - \tilde{A}_i \EE[\sbb|\zb]\|^2] + \| \diag(\VV[\sbb | \zb]) \|^2)
  \end{align}
  are replaced by updates
  \begin{align}
    \sigma_i^2 \leftarrow \frac1{v} (\EE[\| \zb_i - \tilde{A}_i \EE[\sbb|\zb]\|^2] + \| \diag(\VV[\sbb | \zb]) \|^2)
  \end{align}
  so that $\tilde{A}_i = A'_i$.
  %

  Therefore, the updates in both algorithms are linked  by $A_i = U_{\xb_i}A_i'$,
  $\sigma_i' =
  \sigma_i$ and $\Sigma_s'  = \Sigma_s$.
  %

  This concludes the proof.
  %

\end{proof}

Proposition~\ref{prop:optimaldetsrm} and Proposition~\ref{prop:optimalprobsrm} show that
no information is lost by replacing $\xb_i \in \RR^v$ by its reduced
representation $\zb_i \in \RR^n$.
%
A key observation is that Proposition~\ref{prop:optimaldetsrm} and
Proposition~\ref{prop:optimalprobsrm} hold whether or
not the model for deterministic (respectively probabilistic) SRM is indeed the
generative model of the data.
%

A complexity analysis shows that finding the optimal atlas becomes the limiting
step of the pipeline.
%
Even with fast implementations, the subject specific PCA
is costly.
%
However FastSRM only works on $\zb_i$ so we do not need to know the value of
$U_{\xb_i}$.
%
In practice, we observe data $X_i \in \RR^{v \times n}$ and we want to get $Z_i
\in \RR^{n \times n}$ such that $X_i = U_{\xb_i} Z_i$.
%
This can be done by performing an
SVD of $X_i^{\top} X_i$ yielding $X_i^{\top}X_i= V_i D_i V_i^{\top}$ and setting
$Z_i = D_i^{\frac12} V_i^{\top}$.
%
Although computing the product $X_i^{\top} X_i$ has time complexity
$\bigO(vt^2)$ (the same as a PCA) the constant in the $\bigO$ is exactly one so
this operation costs a lot less than the PCA
on full data.
%
When estimates of the mixing matrices are needed, they can be obtained by
applying equation~\eqref{eq:fastdetsrm:Aiupdate} in the deterministic SRM case and
equation~\eqref{eq:fastprobsrm:Aiupdate} in the probabilistic SRM case which only costs
$\bigO(mvp^2)$.
%
In practice the cost of the matrix products $X_i^{\top} X_i$ is often still the
limiting step of the pipeline (this depends on the number of iterations) but as
we show in the next chapter, it is much more efficient than performing SRM on
the full data.
%
Note than if memory allows it, these matrix products can be
computed in parallel.
%

\section{Related work}
The implementation of SRM proposed in~\cite{chen2015reduced} is quadratic in the
number of features which prevent its application to full brain data.
In~\cite{anderson2016enabling}, the authors address this issue by using the
inversion lemma to remove the quadratic dependence. Their implementation is now the most widely used one.
In our work, we propose to reduce further the computation time and memory usage
compared to the implementation in~\cite{anderson2016enabling} by the use of
optimal atlases.

Other popular methods used to speed up SRM are searchlight~\cite{zhang2016searchlight} or piecewise~\cite{bazeille2021empirical}
approaches. While these methods are efficient, they do not optimize the same
objective as the original SRM algorithm and are arguably less principled since the
searchlights or pieces are chosen a priori.

We see SRM as a well principled formulation of the hyperalignment
method~\cite{haxby2011common}.
%
Many methods exist to perform hyperalignment: deep
hyperalignment~\cite{yousefnezhad2017deep} , robust
SRM~\cite{turek2018capturing} , multi-view canonical correlation
analysis~\cite{li2009joint}, ShICA~\cite{richard2021model},
MultiViewICA~\cite{richard2020modeling} optimal
transport~\cite{bazeille2019local} and many more.
%
In this work, we do not claim that FastSRM is the best method. We only see that
SRM is widely used and aim to provide a faster, more memory efficient and
identifiable implementation.  In our view FastSRM should be seen as a plugin
replacement for SRM.

\section{Experiments}
We make several experiments on both synthetic and real data.
%
We used nilearn~\cite{abraham2014machine} for fMRI data preprocessing,
Brainiak~\cite{kumar2020brainiak2} for the non-identifiable version of SRM that
implements the work of~\cite{chen2015reduced} and~\cite{anderson2016enabling}, Numpy~\cite{harris2020array} for
array processing, Scipy~\cite{2020SciPy-NMeth} its implementation of the
Hungarian algorithm, Matplotlib~\cite{hunter2007matplotlib} for plotting and
Sklearn~\cite{pedregosa2011scikit} for machine learning pipelines.

\subsection{Comparing Fitting time and performance of FastSRM and
  SRM on synthetic data}
We generate synthetic data $\xb_i$ according to the model of probabilistic SRM.
%
The parameters $\sigma_i$, $A_i$ and $\Sigma_s$ are generated randomly.
%
We sample the value of the subject specific noise level from a normal
distribution: $\sigma_i \sim \Ncal(0, 0.1)$.
%
The mixing matrices $A_i$
are obtained by sampling their coefficient from a standardized normal
distribution.
%
Lastly, the covariance of the shared response $\Sigma_s$ is diagonal and the
diagonal values are obtained by sampling from a Dirichlet distribution with
parameter $(1 \dots 1)$.
%
We set the number of voxels to $v=125~000$, the number of subjects to $m=10$ and
the number of components to $p=50$.
%
We generate $n=1000$ samples.
%

We benchmark deterministic SRM, probabilistic SRM and their FastSRM counterparts in terms of fitting time and performance.
%
Note that in this section, the identifiable implementation of deterministic SRM
and probabilistic SRM described in section~\ref{sec:probabilisticsrm}
and~\ref{sec:deterministicsrm} is used. 
%
Algorithms are designated by the atlas they use and therefore SRM algorithms are refered to
as \emph{None} because no atlas is used and FastSRM algorithms will have the
label \emph{Optimal}.
%
Note that it would be possible to use FastSRM with sub-optimal
atlases (there exists a wide variety of atlases
available~\cite{schaefer2017local, bellec2010multi, mensch2018extracting}) but
without any guarantees that the performance are the same as SRM.
%

We use a number of iterations between 1 and 100 and report the performance,
fitting time and a measure of convergence.
%
In FastSRM, we do not compute the
unmixing matrices but only the shared response.
%
We measure the performance of an algorithm by computing the error between the true component $S \in \RR^{p \times n}$ and
the predicted component $\hat{S} \in \RR^{p \times n}$ using the quantity:
\begin{align}
\mathrm{mse}(\hat{S}, S) = min_{A \in \RR^{p \times p}} \|A \hat{S} - S \|^2_F =  \|S
\hat{S}^{\dagger}\hat{S} - S \|^2_F
\end{align}
%
This way of measuring errors is
insensitive to the indeterminacies in DetSRM.
%
We measure the fitting time in seconds.
%
Lastly, we measure convergence by computing the gradient $\ell_{\infty}$ norm in
case of DetSRM given by $\max(\mathrm{abs}(G))$ where $G$ is the gradient and
use the distance between consecutive values of the loss for ProbSRM.
%
Results are plotted in Figure~\ref{fig:srm:synthetic_gradient}.
%

We empirically see that the optimal approach is equivalent to using no atlas in
terms of performance.
%
This is predicted by the theory in section~\ref{sec:optimal_atlas} where we demonstrate that these two algorithms yield exactly the same output
from the same input.
%
% %
% Among approximate methods, the projection
% method yields the best results but are quite far away from optimal results.
%
% %
In general probabilistic methods give much better results than their deterministic
counterpart.
%
This shows the superiority of likelihood based methods.
%
In terms of fitting time, FastSRM is about a thousand time faster than SRM after
100 iterations.
%
When no atlas is used, the number
of iterations has a very strong impact on performance while it has a small impact
when the optimal atlas is used.
%
Lastly, looking at the convergence curves, we see that even after $100$ iterations, algorithms did
not fully converge.
%
This means that in practice a much larger number of
iterations is needed which would yield an even higher difference in fitting time
between methods using no atlas and methods using the optimal atlas.
%

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figures/srm/synthetic_gradient.pdf}
  \caption{\textbf{Benchmark of SRM algorithms on synthetic data: } Performance,
    fitting time and convergence of SRM algorithms in the deterministic (left)
    or probabilistic (right) case.
    %
    As expected, when optimal atlases are used,
    the performance is the same as if no atlas is used but the fitting time is
    much lower.
    %
    This is even more pronounced when the number of iterations is
    high (and looking at convergence curves, we see that more iterations could
    be performed to be closer to a stationary point).
    %
  }
  \label{fig:srm:synthetic_gradient}
\end{figure}

\subsection{Experiment on fMRI data: identifiability increases stability}
We evaluate the performance of the different SRM implementations on the
\emph{Sherlock} datasets where fMRI of 17 participants watching "Sherlock" BBC
TV show (episode 1) is performed.
%
% 
These data are downloaded from
\url{http://arks.princeton.edu/ark:/88435/dsp01nz8062179}.
%
% 
Data were acquired using a 3T scanner with an isotropic spatial resolution of
3~mm.
%
% 
More information including the preprocessing pipeline is available in
\cite{sherlock}.
%
% 
Subject 5 is removed because of missing data, leaving us with 16 participants.
%
% 
Although SHERLOCK data contains originally only 1 run, we split it into 4 blocks
of 395 timeframes and one block of 396 timeframes for the needs of our
experiments.
%

We first show that identifiability is a desirable property as it increases
stability.
%
Then we show that our FastSRM algorithm (that works with optimal atlases)
matches the performance of SRM (that works on full data) but use much less
computational resources.
%

\subsection{Identifiability increases stability}
Assuming that the data follow the model, identifiability ensures that the
correct parameters can be recovered without ambiguity.
%
In real fMRI data, the model cannot be expected to hold perfectly, but we can
hope for greater stability in the parameters recovered than if an unidentifiable
model is used.
%

To measure the stability of the common components obtained from the Sherlock
dataset, we divide the subjects into two roughly equal groups and extract the
common components in each group.
%
The components are then matched using the Hungarian algorithm and the stability
measure is obtained from the average correlation of the matched components.
%
The procedure is repeated 9 times using the Brainiak implementation (that is not
identifiable) and our
FastSRM implementation.
%
We plot the results as a scatter plot in Figure~\ref{exp:identifiability}, where, for each repetition , the x-axis indicates the stability measure obtained with Brainiak's implementation
(not identifiable since the shared components covariance is unconstrained) and the y-axis the stability measure
obtained with FastSRM (identifiable since the shared components covariance is diagonal).
%
We see that introducing a diagonal source covariance improves stability.
%

\begin{figure}
\begin{minipage}{0.38\linewidth}
      \includegraphics[width=\textwidth]{figures/srm/identifiability.pdf}
    \end{minipage}
    \hfill
    \begin{minipage}{0.6\linewidth}
        \caption{\label{exp:identifiability}\textbf{Identifiability increases
            stability: } We first divide the subjects
          into two groups and extract the common components in each group.
          %
          The components of the two groups are then matched using the Hungarian
          algorithm and the stability index is determined by the average
          correlation of the matched components.
          %
          The procedure is repeated 9 times with the Brainiak implementation
          (not identifiable since the shared components covariance is unconstrained) and our FastSRM implementation
          (identifiable since the shared components covariance is diagonal).
          %
        }
    \end{minipage}
\end{figure}

\subsection{Comparing fitting time, memory usage and performance on a
  timesegment matching experiment}
\label{sec:timesegment_expe}
\label{timesegment_expe}
The timesegment matching experiment is first introduced
in~\cite{chen2015reduced}.
%
In a nutshell, the time-segment matching accuracy measures the similarity
between two multivariate time-series by trying to localize a time-segment in one
time-series by correlation with the other.
%
In the context of movie watching, this measure has a lot of sense: if we split
the movies in scenes and compute a representation per scene and per subject, it makes
sense to assume that different subjects watching the movie would still have
closer representation of the same scenes than of different scenes.
%
This explains why timesegment matching is a standard evaluation of SRM-like methods also used in  \cite{guntupalli2018computational}, \cite{Nastase741975} or
\cite{zhang2016searchlight}.
%

We now describe more precisely the experimental design.
%
We split the runs into a train and test set.
%
After fitting the model on the
training set, we apply the unmixing matrices $W_i=A_i^{-1}$ of each subject on
the test set yielding individual components matrices.
%
We estimate the shared responses by averaging the individual components of each
subjects but one.
%
We select a target time-segment (9 consecutive timeframes) in the shared
responses and try to localize the corresponding time segment in the components
of the left-out subject using a maximum-correlation classifier.
%

The time-segment is said to be
correctly classified if the correlation between the sample and target
time-segment is higher than with any other time-segment (partially overlapping
time windows are excluded).
%

% 
We use 5-Fold cross-validation across runs: the training set contains 80\% of
the runs and the test set 20\%, and repeat the experiment using all possible
choices for left-out subjects.
%
% 
The mean accuracy is reported in Figure~\ref{fig:srm:timesegment} (bottom).
%
When the
optimal atlas is used, the accuracy is the same as when no atlas is used but
the fitting time is reduced by a factor 10 to 100 and so is the memory usage.
%

%

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{figures/srm/timesegment_matching_sherlock.pdf}
  \caption{\textbf{Benchmark of SRM algorithms on fMRI data} (top) Timesegment
    matching accuracy (middle) Fitting time (bottom) Memory usage. When the
    optimal atlas is used, the accuracy is the same as when no atlas is used but
    the fitting time is reduced by a factor 10 to 100 and so is the memory usage.
    %
  }
  \label{fig:srm:timesegment}
\end{figure}

\section{Conclusion}
As studies using naturalistic stimuli will tend to become more common and large
within and across subjects, we need scalable models in terms of computation time
and memory usage.
%
This is what FastSRM provides.

%
FastSRM is an efficient implementation of SRM
that uses optimal atlases to speed up computations and reduce memory requirements
with provably no loss of performance.

We show on synthetic and real data that FastSRM is much faster and more memory
efficient that implementations not using the optimal atlas.
Furthermore, FastSRM is identifiable. On real data, we show that compared to
Brainiak's implementation (which is not identifiable), FastSRM provides more
stable estimates of the shared components.

%
FastSRM inherits from all the weaknesses of SRM including the fact that
mixing matrices are assumed to be orthogonal which is rather unrealistic. It
remains to be seen whether dimension reduction and learning of the shared
components could be done jointly and efficiently without assuming orthogonal
mixing matrices.
%
Our optimal atlas provide the intuition that SRM-like methods do not need the
full data to provide accurate estimates of shared components. We believe such
insights may guide the design of future methods.

\bibliographystyle{plain}
\bibliography{utils/biblio}
\end{document}

### 2021-11-04 experiments/synthetic.py
update add results

@@ -56,6 +56,7 @@ def do_expe(it, seed, algo):
 
 seeds = np.arange(30)
 iters = 100
+os.makedirs("./results", exist_ok=True)
 for algo in ["detsrm", "probsrm", "fastsrm"]:
     res = Parallel(n_jobs=3, verbose=10)(
         delayed(do_expe)(iters, seed, algo) for seed in seeds

### 2021-11-05 experiments/synthetic.py
update synthetic

@@ -42,7 +42,7 @@ def do_expe(it, seed, algo):
         S = probsrm(
             X, k, n_iter=it, random_state=rng, callback=callback, tol=-1,
         )[-1]
-    if "fastsrm" in algo:
+    if algo == "fastdet":
         S = fastsrm(
             [[x] for x in X],
             k,
@@ -50,6 +50,17 @@ def do_expe(it, seed, algo):
             random_state=rng,
             callback=callback,
             tol=-1,
+            method="det",
+        )[-1]
+    if algo == "fastprob":
+        S = fastsrm(
+            [[x] for x in X],
+            k,
+            n_iter=it,
+            random_state=rng,
+            callback=callback,
+            tol=-1,
+            method="prob",
         )[-1]
     return np.array(S)
 

### 2021-11-05 experiments/synthetic.py
update synthetic

@@ -68,8 +68,8 @@ def do_expe(it, seed, algo):
 seeds = np.arange(30)
 iters = 100
 os.makedirs("./results", exist_ok=True)
-for algo in ["detsrm", "probsrm", "fastsrm"]:
-    res = Parallel(n_jobs=3, verbose=10)(
+for algo in ["detsrm", "probsrm", "fastdet", "fastprob"]:
+    res = Parallel(n_jobs=4, verbose=10)(
         delayed(do_expe)(iters, seed, algo) for seed in seeds
     )
     res = np.array(res)

### 2021-11-05 experiments/synthetic.py
update

@@ -69,7 +69,7 @@ seeds = np.arange(30)
 iters = 100
 os.makedirs("./results", exist_ok=True)
 for algo in ["detsrm", "probsrm", "fastdet", "fastprob"]:
-    res = Parallel(n_jobs=4, verbose=10)(
+    res = Parallel(n_jobs=2, verbose=10)(
         delayed(do_expe)(iters, seed, algo) for seed in seeds
     )
     res = np.array(res)

### 2021-11-05 experiments/synthetic.py
same same

@@ -69,7 +69,7 @@ seeds = np.arange(30)
 iters = 100
 os.makedirs("./results", exist_ok=True)
 for algo in ["detsrm", "probsrm", "fastdet", "fastprob"]:
-    res = Parallel(n_jobs=2, verbose=10)(
+    res = Parallel(n_jobs=4, verbose=10)(
         delayed(do_expe)(iters, seed, algo) for seed in seeds
     )
     res = np.array(res)

### 2021-11-05 experiments/synthetic.py
update

@@ -68,7 +68,8 @@ def do_expe(it, seed, algo):
 seeds = np.arange(30)
 iters = 100
 os.makedirs("./results", exist_ok=True)
-for algo in ["detsrm", "probsrm", "fastdet", "fastprob"]:
+# for algo in ["detsrm", "probsrm", "fastdet", "fastprob"]:
+for algo in ["fastdet"]:
     res = Parallel(n_jobs=4, verbose=10)(
         delayed(do_expe)(iters, seed, algo) for seed in seeds
     )

### 2021-11-05 experiments/synthetic.py
update identifiability

@@ -6,8 +6,6 @@ from fastsrm.utils import reg_error
 import numpy as np
 from joblib import delayed, Parallel
 
-# from brainiak.funcalign.srm import SRM, DetSRM
-
 
 dim = (50, 50, 50)
 m, v, k, n = 10, np.prod(dim), 50, 1000
@@ -68,8 +66,7 @@ def do_expe(it, seed, algo):
 seeds = np.arange(30)
 iters = 100
 os.makedirs("./results", exist_ok=True)
-# for algo in ["detsrm", "probsrm", "fastdet", "fastprob"]:
-for algo in ["fastdet"]:
+for algo in ["detsrm", "probsrm", "fastdet", "fastprob"]:
     res = Parallel(n_jobs=4, verbose=10)(
         delayed(do_expe)(iters, seed, algo) for seed in seeds
     )

### 2021-11-05 plotting/parameters.py
add plotting function

@@ -1,0 +1,24 @@
+NAMES = {
+    "probsrm": "ProbSRM",
+    "detsrm": "DetSRM",
+}
+
+
+def add(NAMES, algo, atlas_only=False, display_regions=True):
+    dic = {}
+    for a in algo:
+        if "fastsrm" not in a:
+            continue
+        method_name, prob_or_det, atlas_name, n_regions = a.split("_")
+        print(prob_or_det)
+        name = atlas_name.upper()
+        if display_regions:
+            name = name + " " + n_regions
+        if not atlas_only:
+            if prob_or_det == "prob":
+                dic[a] = "ProbSRM (" + name + ")"
+            else:
+                dic[a] = "DetSRM (" + name + ")"
+        else:
+            dic[a] = name
+    return {**NAMES, **dic}

### 2021-11-05 plotting/plot_synthetic_gradients.py

@@ -1,0 +1,130 @@
+import numpy as np
+import matplotlib.ticker as plticker
+import matplotlib.pyplot as plt
+from matplotlib.cm import get_cmap
+from parameters import NAMES, add
+import os
+
+os.makedirs("../figures", exist_ok=True)
+loc = plticker.MultipleLocator(base=10)
+
+rc = {
+    "pdf.fonttype": 42,
+    "text.usetex": True,
+    "font.size": 16,
+    "xtick.labelsize": 16,
+    "ytick.labelsize": 16,
+    "text.latex.preview": True,
+    "font.family": "serif",
+}
+plt.rcParams.update(rc)
+
+dim = (50, 50, 50)
+m, v, k, n = 10, np.prod(dim), 50, 1000
+
+algos = ["probsrm", "detsrm"]
+for method in ["prob", "det"]:
+    algos.append("fastsrm_%s_%s_%s" % (method, "rena", "1n"))
+    algos.append("fastsrm_%s_%s_%s" % (method, "proj", "5n"))
+    algos.append("fastsrm_%s_%s_%i" % (method, "pca", k + 1))
+    algos.append("fastsrm_%s_%s_%s" % (method, "pca", "2n"))
+
+
+algos_prob = ["probsrm"]
+for method in ["prob"]:
+    algos_prob.append("fastsrm_%s_%s_%s" % (method, "pca", "2n"))
+
+
+algos_det = ["detsrm"]
+for method in ["det"]:
+    algos_det.append("fastsrm_%s_%s_%s" % (method, "pca", "2n"))
+
+NAMES = add(NAMES, algos_det + algos_prob, atlas_only=True, display_regions=False)
+print(NAMES)
+NAMES["probsrm"] = "None"
+NAMES["detsrm"] = "None"
+NAMES["fastsrm_det_pca_2n"] = "Optimal"
+NAMES["fastsrm_prob_pca_2n"] = "Optimal"
+seeds = np.arange(30)
+
+
+vir = get_cmap("Set2", len(algos_det))
+f, axes = plt.subplots(3, 2, sharex="col", sharey="row", figsize=(10, 9))
+for j, algos_group in enumerate([algos_det, algos_prob]):
+    for c, algo in enumerate(algos_group):
+        res = np.load("../experiments/results/synthetic_grad_%s.npy" % algo)
+        iters = np.arange(len(res[0]))
+        times = res[:, :, 1].astype(float)
+        errors = res[:, :, 0].astype(float)
[…]

### 2021-11-05 examples/synthetic_toy.py
add toy example

@@ -1,0 +1,71 @@
+import os
+from time import time
+from fastsrm.fastsrm import fastsrm
+from fastsrm.srm import detsrm, probsrm, projection
+from fastsrm.utils import reg_error
+import numpy as np
+from joblib import delayed, Parallel
+
+
+dim = (50, 50, 50)
+m, v, k, n = 10, np.prod(dim), 50, 1000
+
+
+def do_expe(it, seed, algo):
+    rng = np.random.RandomState(seed)
+    Sigma = rng.dirichlet(np.ones(k), 1).flatten()
+    S = np.sqrt(Sigma)[:, None] * rng.randn(k, n)
+    W = np.array([projection(rng.randn(v, k)) for i in range(m)])
+    sigmas = 0.1 * rng.rand(m)
+    N = np.array([sigmas[i] * rng.randn(v, n) for i in range(m)])
+    X = np.array([W[i].dot(S) + N[i] for i in range(m)])
+    S_true = S
+    t_init = time()
+
+    def callback(source, gnorm, current_iter, current_time):
+        return (
+            float(reg_error(np.copy(S_true), np.copy(source))),
+            float(current_time - t_init),
+            current_iter,
+            seed,
+            algo,
+            float(gnorm),
+        )
+
+    if algo == "detsrm":
+        S = detsrm(
+            X, k, n_iter=it, random_state=rng, callback=callback, tol=-1,
+        )[-1]
+    if algo == "probsrm":
+        S = probsrm(
+            X, k, n_iter=it, random_state=rng, callback=callback, tol=-1,
+        )[-1]
+    if algo == "fastdet":
+        S = fastsrm(
+            [[x] for x in X],
+            k,
+            n_iter=it,
+            random_state=rng,
+            callback=callback,
+            tol=-1,
+            method="det",
+        )[-1]
+    if algo == "fastprob":
+        S = fastsrm(
+            [[x] for x in X],
+            k,
+            n_iter=it,
+            random_state=rng,
+            callback=callback,
+            tol=-1,
+            method="prob",
+        )[-1]
+    return np.array(S)
+
+
+iters = 20
+os.makedirs("./results", exist_ok=True)
+for algo in ["detsrm", "probsrm", "fastdet", "fastprob"]:
+    S = do_expe(20, 0, algo)
+
+

### 2021-11-05 examples/synthetic_toy.py
update fastdet

@@ -4,11 +4,14 @@ from fastsrm.fastsrm import fastsrm
 from fastsrm.srm import detsrm, probsrm, projection
 from fastsrm.utils import reg_error
 import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.ticker as plticker
+from matplotlib.cm import get_cmap
 from joblib import delayed, Parallel
 
 
-dim = (50, 50, 50)
-m, v, k, n = 10, np.prod(dim), 50, 1000
+dim = (10, 10, 10)
+m, v, k, n = 10, np.prod(dim), 20, 100
 
 
 def do_expe(it, seed, algo):
@@ -65,7 +68,118 @@ def do_expe(it, seed, algo):
 
 iters = 20
 os.makedirs("./results", exist_ok=True)
-for algo in ["detsrm", "probsrm", "fastdet", "fastprob"]:
-    S = do_expe(20, 0, algo)
+algos_det = ["detsrm", "fastdet"]
+algos_prob = ["probsrm", "fastprob"]
+algos = algos_det + algos_prob
+S = []
+for algo in algos:
+    resS = do_expe(20, 0, algo)
+    print(algo, "len", len(resS), len(resS[0]))
+    print(resS[0])
+    S.append(resS)
+
+index = {a: k for k, a in enumerate(algos)}
+loc = plticker.MultipleLocator(base=10)
+
+rc = {
+    "pdf.fonttype": 42,
+    "text.usetex": True,
+    "font.size": 16,
+    "xtick.labelsize": 16,
+    "ytick.labelsize": 16,
+    "text.latex.preview": True,
+    "font.family": "serif",
+}
+plt.rcParams.update(rc)
+
+NAMES = {}
+NAMES["probsrm"] = "None"
+NAMES["detsrm"] = "None"
+NAMES["fastdet"] = "Optimal"
+NAMES["fastprob"] = "Optimal"
+seeds = np.arange(30)
 
 
+vir = get_cmap("Set2", len(algos_det))
+f, axes = plt.subplots(3, 2, sharex="col", sharey="row", figsize=(10, 9))
+for j, algos_group in enumerate([algos_det, algos_prob]):
+    for c, algo in enumerate(algos_group):
+        res = S[index[algo]]
+        iters = np.arange(len(res[0]))
+        times = res[:, :, 1].astype(float)
+        errors = res[:, :, 0].astype(float)
+        gnorms = res[:, :, 5].astype(float)
+        linestyle = "-"
[…]

### 2021-11-05 fastsrm/fastsrm.py

@@ -205,7 +205,13 @@ errors when the number of subjects and/or sessions is large
 
     if method == "det":
         ressrm = detsrm(
-            Xred, n_components, n_iter, random_state, verbose=verbose, tol=tol,
+            Xred,
+            n_components,
+            n_iter,
+            random_state,
+            verbose=verbose,
+            tol=tol,
+            callback=callback,
         )
 
     S = ressrm[1]

### 2021-11-05 experiments/synthetic.py
redo fastdet

@@ -66,7 +66,8 @@ def do_expe(it, seed, algo):
 seeds = np.arange(30)
 iters = 100
 os.makedirs("./results", exist_ok=True)
-for algo in ["detsrm", "probsrm", "fastdet", "fastprob"]:
+# for algo in ["detsrm", "probsrm", "fastdet", "fastprob"]:
+for algo in ["fastdet"]:
     res = Parallel(n_jobs=4, verbose=10)(
         delayed(do_expe)(iters, seed, algo) for seed in seeds
     )

### 2021-11-05 examples/plot_synthetic_toy.py
plot toy example

@@ -1,3 +1,12 @@
+"""
+============================================================
+FastSRM : running time and convergence
+============================================================
+
+This example demonstrates on a very small dataset that
+the optimal atlas used in FastSRM speeds up the computation
+of shared components but without any loss of accuracy.
+"""
 import os
 from time import time
 from fastsrm.fastsrm import fastsrm

### 2021-11-05 plotting/parameters.py
update README

@@ -2,23 +2,3 @@ NAMES = {
     "probsrm": "ProbSRM",
     "detsrm": "DetSRM",
 }
-
-
-def add(NAMES, algo, atlas_only=False, display_regions=True):
-    dic = {}
-    for a in algo:
-        if "fastsrm" not in a:
-            continue
-        method_name, prob_or_det, atlas_name, n_regions = a.split("_")
-        print(prob_or_det)
-        name = atlas_name.upper()
-        if display_regions:
-            name = name + " " + n_regions
-        if not atlas_only:
-            if prob_or_det == "prob":
-                dic[a] = "ProbSRM (" + name + ")"
-            else:
-                dic[a] = "DetSRM (" + name + ")"
-        else:
-            dic[a] = name
-    return {**NAMES, **dic}

### 2021-11-05 plotting/plot_synthetic_gradients.py

@@ -2,7 +2,7 @@ import numpy as np
 import matplotlib.ticker as plticker
 import matplotlib.pyplot as plt
 from matplotlib.cm import get_cmap
-from parameters import NAMES, add
+from parameters import NAMES
 import os
 
 os.makedirs("../figures", exist_ok=True)
@@ -22,29 +22,15 @@ plt.rcParams.update(rc)
 dim = (50, 50, 50)
 m, v, k, n = 10, np.prod(dim), 50, 1000
 
-algos = ["probsrm", "detsrm"]
-for method in ["prob", "det"]:
-    algos.append("fastsrm_%s_%s_%s" % (method, "rena", "1n"))
-    algos.append("fastsrm_%s_%s_%s" % (method, "proj", "5n"))
-    algos.append("fastsrm_%s_%s_%i" % (method, "pca", k + 1))
-    algos.append("fastsrm_%s_%s_%s" % (method, "pca", "2n"))
+algos_det = ["detsrm", "fastdet"]
+algos_prob = ["probsrm", "fastprob"]
 
-
-algos_prob = ["probsrm"]
-for method in ["prob"]:
-    algos_prob.append("fastsrm_%s_%s_%s" % (method, "pca", "2n"))
-
-
-algos_det = ["detsrm"]
-for method in ["det"]:
-    algos_det.append("fastsrm_%s_%s_%s" % (method, "pca", "2n"))
-
-NAMES = add(NAMES, algos_det + algos_prob, atlas_only=True, display_regions=False)
-print(NAMES)
+NAMES = {}
 NAMES["probsrm"] = "None"
 NAMES["detsrm"] = "None"
-NAMES["fastsrm_det_pca_2n"] = "Optimal"
-NAMES["fastsrm_prob_pca_2n"] = "Optimal"
+NAMES["fastdet"] = "Optimal"
+NAMES["fastprob"] = "Optimal"
+
 seeds = np.arange(30)
 
 

### 2021-11-05 examples/plot_synthetic_toy.py
add synthetic plot

@@ -20,7 +20,7 @@ from joblib import delayed, Parallel
 
 
 dim = (15, 15, 15)
-m, v, k, n = 10, np.prod(dim), 20, 100
+m, v, k, n = 5, np.prod(dim), 20, 30
 
 
 def do_expe(it, seed, algo):
@@ -175,7 +175,6 @@ axes[2, 0].set_yscale("log")
 y_minor = plticker.LogLocator(
     base=10.0, subs=np.arange(1.0, 10.0) * 0.1, numticks=10
 )
-axes[1, 0].set_ylim([1e-2, 1e0])
 axes[1, 0].set_yticks([1e-2, 1e-1, 1e0])
 axes[0, 0].yaxis.set_minor_locator(y_minor)
 axes[1, 0].yaxis.set_minor_locator(y_minor)
@@ -189,3 +188,4 @@ axes[2, 1].set_xlabel("Number of iterations")
 plt.tick_params(axis="y", which="minor")
 plt.legend(loc="upper center", bbox_to_anchor=(-0.1, 4), ncol=5, title="ATLAS")
 plt.show()
+plt.savefig("./fig.pdf", bbox_inches="tight")

### 2021-11-05 examples/plot_synthetic_toy.py
remove save fig

@@ -188,4 +188,3 @@ axes[2, 1].set_xlabel("Number of iterations")
 plt.tick_params(axis="y", which="minor")
 plt.legend(loc="upper center", bbox_to_anchor=(-0.1, 4), ncol=5, title="ATLAS")
 plt.show()
-plt.savefig("./fig.pdf", bbox_inches="tight")

### 2021-11-05 examples/plot_synthetic_toy.py
add print

@@ -188,3 +188,4 @@ axes[2, 1].set_xlabel("Number of iterations")
 plt.tick_params(axis="y", which="minor")
 plt.legend(loc="upper center", bbox_to_anchor=(-0.1, 4), ncol=5, title="ATLAS")
 plt.show()
+print("Done")

### 2021-11-05 examples/plot_synthetic_toy.py
update

@@ -187,5 +187,6 @@ axes[2, 0].set_xlabel("Number of iterations")
 axes[2, 1].set_xlabel("Number of iterations")
 plt.tick_params(axis="y", which="minor")
 plt.legend(loc="upper center", bbox_to_anchor=(-0.1, 4), ncol=5, title="ATLAS")
+f.tight_layout()
 plt.show()
 print("Done")

### 2021-11-05 examples/plot_synthetic_toy.py
add verbose

@@ -17,6 +17,7 @@ import matplotlib.pyplot as plt
 import matplotlib.ticker as plticker
 from matplotlib.cm import get_cmap
 from joblib import delayed, Parallel
+print("Starting FastSRM example")
 
 
 dim = (15, 15, 15)
@@ -187,6 +188,5 @@ axes[2, 0].set_xlabel("Number of iterations")
 axes[2, 1].set_xlabel("Number of iterations")
 plt.tick_params(axis="y", which="minor")
 plt.legend(loc="upper center", bbox_to_anchor=(-0.1, 4), ncol=5, title="ATLAS")
-f.tight_layout()
 plt.show()
 print("Done")

### 2021-11-05 examples/plot_synthetic_toy.py
update example

@@ -89,17 +89,6 @@ for algo in algos:
 index = {a: k for k, a in enumerate(algos)}
 loc = plticker.MultipleLocator(base=10)
 
-rc = {
-    "pdf.fonttype": 42,
-    "text.usetex": True,
-    "font.size": 16,
-    "xtick.labelsize": 16,
-    "ytick.labelsize": 16,
-    "text.latex.preview": True,
-    "font.family": "serif",
-}
-plt.rcParams.update(rc)
-
 NAMES = {}
 NAMES["probsrm"] = "None"
 NAMES["detsrm"] = "None"

### 2021-11-05 _downloads/eefa709bfa15f19e4cbfadab95d67431/plot_synthetic_toy.py
[skip ci] Updates

@@ -1,0 +1,181 @@
+"""
+============================================================
+FastSRM : running time and convergence
+============================================================
+
+This example demonstrates on a very small dataset that
+the optimal atlas used in FastSRM speeds up the computation
+of shared components but without any loss of accuracy.
+"""
+import os
+from time import time
+from fastsrm.fastsrm import fastsrm
+from fastsrm.srm import detsrm, probsrm, projection
+from fastsrm.utils import reg_error
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.ticker as plticker
+from matplotlib.cm import get_cmap
+from joblib import delayed, Parallel
+print("Starting FastSRM example")
+
+
+dim = (15, 15, 15)
+m, v, k, n = 5, np.prod(dim), 20, 30
+
+
+def do_expe(it, seed, algo):
+    rng = np.random.RandomState(seed)
+    Sigma = rng.dirichlet(np.ones(k), 1).flatten()
+    S = np.sqrt(Sigma)[:, None] * rng.randn(k, n)
+    W = np.array([projection(rng.randn(v, k)) for i in range(m)])
+    sigmas = 0.1 * rng.rand(m)
+    N = np.array([sigmas[i] * rng.randn(v, n) for i in range(m)])
+    X = np.array([W[i].dot(S) + N[i] for i in range(m)])
+    S_true = S
+    t_init = time()
+
+    def callback(source, gnorm, current_iter, current_time):
+        return (
+            float(reg_error(np.copy(S_true), np.copy(source))),
+            float(current_time - t_init),
+            current_iter,
+            seed,
+            algo,
+            float(gnorm),
+        )
+
+    if algo == "detsrm":
+        S = detsrm(
+            X, k, n_iter=it, random_state=rng, callback=callback, tol=-1,
+        )[-1]
+    if algo == "probsrm":
+        S = probsrm(
+            X, k, n_iter=it, random_state=rng, callback=callback, tol=-1,
+        )[-1]
+    if algo == "fastdet":
+        S = fastsrm(
+            [[x] for x in X],
+            k,
+            n_iter=it,
+            random_state=rng,
+            callback=callback,
+            tol=-1,
+            method="det",
+        )[-1]
+    if algo == "fastprob":
+        S = fastsrm(
[…]

### 2021-11-05 _downloads/eefa709bfa15f19e4cbfadab95d67431/plot_synthetic_toy.py
[skip ci] Updates

@@ -176,6 +176,6 @@ axes[0, 0].set_ylabel("Mean squared error")
 axes[2, 0].set_xlabel("Number of iterations")
 axes[2, 1].set_xlabel("Number of iterations")
 plt.tick_params(axis="y", which="minor")
-plt.legend(loc="upper center", bbox_to_anchor=(-0.1, 4), ncol=5, title="ATLAS")
+plt.legend(loc="upper center", bbox_to_anchor=(-0.1, 3.8), ncol=5, title="ATLAS")
 plt.show()
 print("Done")

### 2021-11-16 plotting/plot_identifiability.py
remove temporary files

@@ -17,8 +17,9 @@ rc = {
 }
 plt.rcParams.update(rc)
 shared_dir = (
-    "/home/hugo/Research/drago/store3/work/hrichard/fastsrm/identifiability/shared/"
+    "../experiments/results/identifiability/shared/"
 )
+
 
 def standardize(X):
     X_ = X - np.mean(X, axis=1)
@@ -49,4 +50,4 @@ plt.xlim(np.min(errors) - 0.1, np.max(errors) + 0.1)
 plt.ylim(np.min(errors) - 0.1, np.max(errors) + 0.1)
 plt.xlabel("Stability index \n (General covariance)")
 plt.ylabel("Stability index \n (Diagonal covariance)")
-plt.savefig("../paper/figures/identifiability.pdf", bbox_inches="tight")
+plt.savefig("../figures/identifiability.pdf", bbox_inches="tight")

### 2021-11-16 plotting/plot_timesegment.py
update plotting and tex

@@ -1,0 +1,93 @@
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+from matplotlib.cm import get_cmap
+
+rc = {
+    "pdf.fonttype": 42,
+    "text.usetex": True,
+    "font.size": 16,
+    "xtick.labelsize": 16,
+    "ytick.labelsize": 16,
+    "text.latex.preview": True,
+    "font.family": "serif",
+}
+plt.rcParams.update(rc)
+
+vir = get_cmap("Set2", 2)
+
+
+algos_prob = ["prob", "fastprob"]
+algos_det = ["det", "fastdet"]
+
+NAMES = {}
+NAMES["prob"] = "None"
+NAMES["det"] = "None"
+NAMES["fastprob"] = "Optimal"
+NAMES["fastdet"] = "Optimal"
+
+sessions = np.arange(5)
+datasets = ["sherlock"]
+n_components = [5, 10, 20, 50]
+for dataset in datasets:
+    vir = get_cmap("Set2", len(algos_det))
+    fig, axes = plt.subplots(3, 2, sharex="col", sharey="row", figsize=(10, 9))
+    plt.suptitle("Dataset: %s" % dataset.upper())
+    for j, algos_group in enumerate([algos_det, algos_prob]):
+        for c, algo in enumerate(algos_group):
+            for i, path_name in enumerate(
+                [
+                    "cv_accuracy_timesegment_matching",
+                    "fit_time_timesegmentmatching",
+                    "memory_usage_timesegmentmatching",
+                ]
+            ):
+                highs = []
+                lows = []
+                meds = []
+                for k in n_components:
+                    datas = []
+                    for session in sessions:
+                        path = "../experiments/results/%s-%i-%i-%s-%s.npy" % (
+                            path_name,
+                            session,
+                            k,
+                            algo,
+                            dataset,
+                        )
+                        data = np.load(path)
+                        datas.append(data)
+                    datas = np.array(datas).flatten()
+                    lows.append(np.quantile(datas, 0.25))
+                    highs.append(np.quantile(datas, 0.75))
+                    meds.append(np.median(datas))
+                if "fast" in algo:
+                    marker = "x"
+                else:
+                    marker = "o"
[…]

### 2021-11-21 paper/main.tex
update plotting and tex
pass by BT

@@
 \usepackage{microtype}      % microtypography
-\usepackage{todonotes}
+% \usepackage{todonotes}
 \usepackage{longtable}
@@
 
+\usepackage[colorinlistoftodos,textwidth=2.1cm]{todonotes}
+\newcommand{\bt}[1]{\todo[color=orange, size=\tiny]{BT: #1}}
+
 \begin{document}
@@
   However when the number
-of subjects or runs is large, fitting the model requires a large amount of memory
+of subjects or runs is large,\bt{True, but we the experiments don't make it so explicit ?} fitting the model requires a large amount of memory
 and computational power, which limits its use in practice.
@@
 not identifiable which makes the shared response difficult to interpret.
-%
 
@@
 than competitive implementations.
-%
 
@@
 \section{Introduction}
-What makes the analysis of functional magnetic resonance imaging (fMRI) data
+\bt{We need to start with a more application-oriented intro}
+%
+There is a solid tradition of multivariate analysis of neuroimaging
+data, that makes it possible to assess the information carried by the
+data even under complex stimulation frameworks, eg movie watching,
+story or music listening \bt{to be continued}.
+%
+%
+What makes the analysis of functional magnetic resonance imaging (fMRI)
+\bt{Question of scope: do want to target only fMRI or only MEEG ? Or only mention MEEG as possible extension in the discussion ?} data
 particularly difficult is that the number of features (voxels) is usually much
-larger than the number of available samples (timeframes).
+larger than the number of available samples (time frames).
 %
-In such cases, dimension reduction methods are precious tools that allow to
+In such cases, dimension reduction methods are needed to
 reduce the computational burden of subsequent data analysis.
@@
 In this work, we introduce FastSRM, an identifiable model that uses an optimal
-atlas based decomposition to speed up the computations with provably no loss of
+atlas-based decomposition to speed up the computations with provably no loss of
 performance.
@@
 estimates of the shared components than its non-identifiable counter-part and 
-on synthetic as well as on real data, it is much faster and more memory
+ it is much faster and more memory
 efficient than other implementations that do not make use of the optimal atlas.
-%
+% on synthetic as well as on real data,
 
@@
   \centering
-  \includegraphics[scale=0.3]{figures/srm/conceptual_figure31.png}
+  \includegraphics[scale=0.3]{figures/srm/conceptual_figure31.png}\bt{in the figure, can you reaplce ``timeframe'' with ``time frame'' ?}
   \caption{\textbf{Shared response model}: The raw fMRI data are modeled as a
@@
 matrices $X_i \in \RR^{v, n}$ and $S \in \RR^{p, n}$.
-%
 
-The deterministic shared response model sees both the mixing matrices $A_i$ and
+The deterministic shared response model considers both the mixing matrices $A_i$ and
 the $n$ observations of the shared response $S$ as parameters to be
@@
 \Sigma_i=\sigma^2 I_v$ where $\sigma$ is an hyper-parameter to choose.
+
+The model is optimized by maximizing the log-likelihood, assuming that
+the noise is Gaussian distributed.
 %
-
-The model is optimized by maximizing the log-likelihood.
-%
-
-The likelihood is given by: $p(\xb) = \prod_i \Ncal(\xb_i; A_i \sbb, \sigma^2 I)$ and
+The likelihood is then given by: $p(\xb) = \prod_i \Ncal(\xb_i; A_i \sbb, \sigma^2 I)$ and
 therefore the empirical expected negative log-likelihood is given up to a constant independent of
@@
 Note that the hyper-parameter $\sigma$ does not have an influence on
-the loss and can therefore be safely ignored.
-%
+the loss and can therefore be ignored.
 
@@
 \begin{align}
-  p(\xb, \sbb) &= \prod_i \frac{\exp(-\frac{\|\xb_i - A_i \sbb \|^2}{2 \sigma_i^2})}{(2 \pi \sigma_i^{2v})^{\frac12}} \frac{\exp(-\frac12 \dotp{ \sbb }{ \Sigma_s^{-1} \sbb } )}{(2 \pi | \Sigma_s|)^{\frac12}} \\
-               &= c_1 \exp(-\frac12 \left( \sum_i \frac1{\sigma_i^2}\|\xb_i\|^2 - 2  \dotp{ \sum_i \frac1{\sigma_i^2} A_i^{\top}\xb_i}{ \sbb } \right.
+  p(\xb, \sbb) &= \prod_i \frac{\exp \left(-\frac{\|\xb_i - A_i \sbb \|^2}{2 \sigma_i^2} \right) }{(2 \pi \sigma_i^{2v})^{\frac12}} \frac{\exp(-\frac12 \dotp{ \sbb }{ \Sigma_s^{-1} \sbb } )}{(2 \pi | \Sigma_s|)^{\frac12}} \\
+               &= c_1 \exp (-\frac12 \left( \sum_i \frac1{\sigma_i^2}\|\xb_i\|^2 - 2  \dotp{ \sum_i \frac1{\sigma_i^2} A_i^{\top}\xb_i}{ \sbb } \right.
 \\& \left.+ \sum_i \frac1{\sigma_i^2} \| \sbb \|^2 + \dotp{ \sbb}{ \Sigma_s^{-1} \sbb }  \right)) \\
@@
 \begin{align}
-  c_2(\xb) = c_1 \exp(-\frac12( \sum_i
+  c_2(\xb) = c_1 \exp \left( -\frac12( \sum_i
   \frac1{\sigma_i^2}\|\xb_i\|^2 - \dotp{  \EE[\sbb }{ \xb], \VV[\sbb | \xb]^{-1}
-  \EE[\sbb | \xb] }))
+  \EE[\sbb | \xb] }) \right)
 \end{align}
-are independent of $\sbb$.
+are independent from $\sbb$.
 %
@@
 \begin{align}
-&\sigma_i^2 \leftarrow \frac1{v} (\EE[\| \xb_i - A_i \EE[\sbb|\xb]\|^2] + \| \diag(\VV[\sbb | \xb]) \|^2) \\
+  &\sigma_i^2 \leftarrow \frac1{v} (\EE[\| \xb_i - A_i \EE[\sbb|\xb]\|^2] + \| \diag(\VV[\sbb | \xb]) \|) \\
+  % no square here, right ?
   &A_i \leftarrow \Pcal(\EE[\xb_i \EE[\sbb|\xb]^{\top}]) \label{eq:psrm:Aiupdate} \\
@@
 %
-In deterministic SRM $\theta = (A_i)_{i=1}^m, S$ where $(A_i)_{i=1}^m$ are the
-mixing matrices and $S$ are the $n$ observations of the shared response $S$ while in probabilistic SRM $\theta
-= (A_i)_{i=1}^m, \Sigma_s, (\sigma_i)_{i=1}^m$ where $(A_i)_{i=1}^m$ are the
+In deterministic SRM $\theta = \left((A_i)_{i=1}^m, S \right)$ where $(A_i)_{i=1}^m$ are the
+mixing matrices and $S$ is the shared response, while in probabilistic SRM $\theta
+= \left( (A_i)_{i=1}^m, \Sigma_s, (\sigma_i)_{i=1}^m \right) $ where $(A_i)_{i=1}^m$ are the
 mixing matrices, $(\sigma_i)_{i=1}^m$ the noise levels and $\Sigma_s$ the
@@
 
-In FastSRM we consider a set of view specific atlases $U_i \in \RR^{v \times r}$ such that
+In FastSRM we consider a set of view-specific atlases $U_i \in \RR^{v \times r}$ such that
 $U_i^{\top}U_i = I_r$ where $r$ is the number of regions in the atlas.
@@
 %
-However, as we will see in the next section, there exists atlases such as the
-correspondence between $\theta$ and
-$\theta'$ can be made explicit.
+However, as we will see in the next section, there exists atlases such that the
+models parametrized by $\theta$ and
+$\theta'$ are equivalent.
 %
@@
 
-From a computational stand point, the dimension reduction provides
+From a computational stand point, dimension reduction provides
 a large reduction in memory usage.
@@
 the memory consumption is only in $\bigO(vn)$ (where $v$ is the number of voxels
-and $n$ is the number of samples) which is lower than SRM by a factor of $m$,
+and $n$ is the number of samples) which is lower than that of SRM by a factor of $m$,
 the number of subjects.
@@
 %
-Let us highlight that an experiment is often run
-multiple times such as when cross validated results are needed.
+Let us highlight that an experiment is often run multiple times,
+typically when the analysis requires cross-validation procedures.
 %
@@
 the number of voxels is about $300~000$ so that $\frac{v}{r}$ is typically about
-$1000$.
+$300$.
 %
-It remains to show how to draw a correspondence between FastSRM and SRM which is
+It remains to be shown how to draw a correspondence between FastSRM and SRM which is
 addressed in the following section.
@@
 sense that SRM and FastSRM yield the same results.
-%
 
-Let us consider $\xb_i = U_{\xb_i} \zb_i$ a PCA of $\xb_i$ with the maximum
+Let us consider $\xb_i = U_{\xb_i} \zb_i$ \bt{But any fill-rank orthogonal basis would work equally well (i.e. any orthogonal transform of the PCA basis)} a PCA of $\xb_i$ with the maximum
 number of components.
@@
 \begin{align}
-  A_i &\leftarrow \Pcal(\frac1{n}X_i S^{\top})= U_{X_i}\Pcal(\frac1{n}Z_i S^{\top})
+  A_i &\leftarrow \Pcal \left(\frac1{n}X_i S^{\top} \right)= U_{X_i}\Pcal \left(\frac1{n}Z_i S^{\top} \right)
   \label{eq:fastdetsrm:Aiupdate}
@@
 
-We call probSRM($\psi$) the probabilistic SRM algorithm modified such that
-updates
+We call probSRM($\psi$)\bt{I don't unserstand what $\psi$ represents}
+the probabilistic SRM algorithm modified such that updates
 \begin{align}
@@
 limiting step of the pipeline (this depends on the number of iterations) but as
-we show in the next chapter, it is much more efficient than performing SRM on
+we show next, it is much more efficient than performing SRM on
 the full data.
 %
-Note than if memory allows it, these matrix products can be
+Note that if memory allows it, these matrix products can be
 computed in parallel.
@@
 %
-In this work, we do not claim that FastSRM is the best method. We only see that
-SRM is widely used and aim to provide a faster, more memory efficient and
-identifiable implementation.  In our view FastSRM should be seen as a plugin
-replacement for SRM.
+In this work, we do not discuss the merit of FastSRM with respect to
+other methods. We only see that SRM is widely used and aim to provide
+a faster, more memory efficient and identifiable implementation.  In
+our view FastSRM should be seen as a plugin replacement for SRM.
 
@@
 Sklearn~\cite{pedregosa2011scikit} for machine learning pipelines.
+\bt{Please provide versions of the software used}
 
@@
 We sample the value of the subject specific noise level from a normal
-distribution: $\sigma_i \sim \Ncal(0, 0.1)$.
+distribution: $\sigma_i \sim \Ncal(0, 0.1)$ \bt{$\sigma_i$ has to be non-negative}.
 %
@@
 Note that it would be possible to use FastSRM with sub-optimal
-atlases (there exists a wide variety of atlases
-available~\cite{schaefer2017local, bellec2010multi, mensch2018extracting}) but
-without any guarantees that the performance are the same as SRM.
+atlases (there exists a wide variety of brain atlases
+~\cite{schaefer2017local, bellec2010multi, mensch2018extracting}) but
+using them does not brain any guarantee that the performance is the same as SRM.
 %
@@
 \begin{align}
-\mathrm{mse}(\hat{S}, S) = min_{A \in \RR^{p \times p}} \|A \hat{S} - S \|^2_F =  \|S
+\mathrm{mse}(\hat{S}, S) = min_{B \in \RR^{p \times p}} \|B \hat{S} - S \|^2_F =  \|S
 \hat{S}^{\dagger}\hat{S} - S \|^2_F
@@
 %
-This means that in practice a much larger number of
-iterations is needed which would yield an even higher difference in fitting time
+This means that in practice a larger number of
+iterations is needed, which would yield an even higher difference in fitting time
 between methods using no atlas and methods using the optimal atlas.
@@
   \centering
-  \includegraphics[width=\textwidth]{figures/srm/synthetic_gradient.pdf}
+  \includegraphics[width=\textwidth]{../figures/synthetic_gradient.pdf}
   \caption{\textbf{Benchmark of SRM algorithms on synthetic data: } Performance,
@@
 Assuming that the data follow the model, identifiability ensures that the
-correct parameters can be recovered without ambiguity.
+correct parameters can be recovered without ambiguity.\bt{I find this a bit ambiguous: you give te impression that you gurantee that the optimal solution is used. Rephrase.}
 %
@@
 \begin{minipage}{0.38\linewidth}
-      \includegraphics[width=\textwidth]{figures/srm/identifiability.pdf}
+      \includegraphics[width=\textwidth]{../figures/identifiability.pdf}
     \end{minipage}
@@
 \label{timesegment_expe}
-The timesegment matching experiment is first introduced
+Time-segment matching experiment is first introduced
 in~\cite{chen2015reduced}.
@@
   \centering
-  \includegraphics[width=0.8\textwidth]{figures/srm/timesegment_matching_sherlock.pdf}
+  \includegraphics[width=0.8\textwidth]{../figures/timesegment_matching_sherlock.pdf}
   \caption{\textbf{Benchmark of SRM algorithms on fMRI data} (top) Timesegment

### 2021-11-21 paper/main.tex
minro stuff

@@
 \label{timesegment_expe}
-Time-segment matching experiment is first introduced
+Time-segment matching has first been introduced
 in~\cite{chen2015reduced}.
@@
 %
-In the context of movie watching, this measure has a lot of sense: if we split
-the movies in scenes and compute a representation per scene and per subject, it makes
-sense to assume that different subjects watching the movie would still have
-closer representation of the same scenes than of different scenes.
+In the context of movie watching, this measure is quite meaningful: if
+we split the movies in scenes and compute a representation per scene
+and per subject, it can be assumed that different subjects watching
+the movie would still have closer representation of the same scenes
+than of different scenes.
 %
@@
 % 
-The mean accuracy is reported in Figure~\ref{fig:srm:timesegment} (bottom).
+The mean accuracy is reported in Figure~\ref{fig:srm:timesegment} (top).
 %
-When the
-optimal atlas is used, the accuracy is the same as when no atlas is used but
-the fitting time is reduced by a factor 10 to 100 and so is the memory usage.
+When the optimal atlas is used, the accuracy is the same as when no
+atlas is used but the fitting time is reduced by a factor 10 to 100
+and so is the memory usage (see Figure~\ref{fig:srm:timesegment},
+bottom).
 %
@@
 %
+This is particularly related to the development of ddep phenotyping models \bt{cite current intensive acquisition endeavours, such as courtois-Neuromod}.
+%
 This is what FastSRM provides.
@@
 %
-Our optimal atlas provide the intuition that SRM-like methods do not need the
+Our optimal atlas apprach shows that SRM-like methods do not need the
 full data to provide accurate estimates of shared components. We believe such
 insights may guide the design of future methods.
+\bt{Add a paragraph on implementation, sofwtare quality, documentation, examples, availability...}
+\bt{Discuss the use for other brain imaging modelities (MEEG) if relevant ?}
 

### 2021-11-25 paper/main.tex
update pytests

@@
 \usepackage[colorinlistoftodos,textwidth=2.1cm]{todonotes}
-\newcommand{\bt}[1]{\todo[color=orange, size=\tiny]{BT: #1}}
+\newcommand{\bt}[1]{\todo[color=orange, inline=True]{BT: #1}}
 
@@
   However when the number
-of subjects or runs is large,\bt{True, but we the experiments don't make it so explicit ?} fitting the model requires a large amount of memory
+  of subjects or runs is large,
+  % \bt{True, but we the experiments don't make it so explicit ?}
+  fitting the model requires a large amount of memory
 and computational power, which limits its use in practice.
@@
 \section{Introduction}
-\bt{We need to start with a more application-oriented intro}
-%
-There is a solid tradition of multivariate analysis of neuroimaging
-data, that makes it possible to assess the information carried by the
-data even under complex stimulation frameworks, eg movie watching,
-story or music listening \bt{to be continued}.
-%
-%
-What makes the analysis of functional magnetic resonance imaging (fMRI)
-\bt{Question of scope: do want to target only fMRI or only MEEG ? Or only mention MEEG as possible extension in the discussion ?} data
-particularly difficult is that the number of features (voxels) is usually much
+%\bt{We need to start with a more application-oriented intro}
+%
+Complex stimulations such as movie-watching, story or music listening are more
+and more popular in the neuro-scientific community. Indeed such naturalistic
+paradigms are unconstrained from behavioral manipulations and thus, more
+ecological with respect to real-world conditions.
+%
+However and in contrast to more traditional designs, these so-called
+naturalistic stimuli do not come with a design matrix that quantifies the
+features of interest in the stimuli.
+%
+Without a design matrix, models such as the general linear
+model~\cite{poline2012general} cannot be used. Although some works have used
+natural annotations~\cite{huth2012continuous} or deep neural
+networks~\cite{gucclu2017increasingly}~\cite{richard2018optimizing} to create an
+estimate of the design matrix of naturalistic
+stimuli it is by no means a simple task.
+
+Another approach is to learn jointly the design matrix and the spatial maps in
+an unsupervised way.
+%
+In the case where the dataset contains only one subject
+independent component analysis~\cite{jutten1991blind} (ICA) is the method of choice.
+%
+ICA assumes that components are independent, a rather realistic assumption which
+ensures identifiability up to permutation and scaling of the model and can be fit efficiently.
+%
+
+But when the dataset contains multiple subjects, a natural assumption is to
+assume that the design matrix is shared across subjects.
+%
+Many different methods
+can produce shared components from different subjects. Some assume independent
+components~\cite{richard2021model}~\cite{richard2020modeling}~\cite{varoquaux2009canica}~\cite{calhoun2001method}.
+while others are only based on second order information such as Multiset
+Canonical Correlation Analysis~\cite{via2011joint} or the Shared Response Model (SRM)~\cite{chen2015reduced}.
+%
+
+This work focuses on the SRM. Because of its
+simplicity and its built-in dimension reduction, this model is widely
+used~\cite{baldassano2017discovering}~\cite{cohen2017computational}~\cite{baldassano2018representation}~\cite{jolly2020custom}~\cite{lee2021anticipation}
+including as a preprocessing step for source
+identification~\cite{richard2021model}~\cite{richard2020modeling},
+classification~\cite{turek2018capturing}~\cite{chen2017shared}\cite{zhang2016searchlight} or as a basis for transfer learning~\cite{zhang2018transfer}.
+
+What makes the analysis of fMRI data particularly difficult is that
+the number of features (voxels) is usually much
 larger than the number of available samples (time frames).
-%
-In such cases, dimension reduction methods are needed to
-reduce the computational burden of subsequent data analysis.
-%
-The shared response
-model~\cite{chen2015reduced} provides a principled way to perform dimension
-reduction of brain imaging data from subjects performing the same task and is
-therefore a natural basis for common source
-identification~\cite{richard2021model}~\cite{richard2020modeling},
-classification~\cite{turek2018capturing}~\cite{chen2017shared}\cite{zhang2016searchlight}
-or transfer learning~\cite{zhang2018transfer}.
-%
-
-However SRM has initially been designed to work within regions of interest using
+Yet, SRM has initially been designed to work within regions of interest using
 only few subjects.
@@
 %
-Another issue of SRM is its non-identifiability (see the proof
+Lastly, another issue of SRM is its non-identifiability (see the proof
 in~\cite{richard2020modeling} Appendix D) which makes it challenging to
@@
   \centering
-  \includegraphics[scale=0.3]{figures/srm/conceptual_figure31.png}\bt{in the figure, can you reaplce ``timeframe'' with ``time frame'' ?}
+  \includegraphics[scale=0.3]{figures/srm/conceptual_figure33.png}
   \caption{\textbf{Shared response model}: The raw fMRI data are modeled as a
@@
 
-Let us consider $\xb_i = U_{\xb_i} \zb_i$ \bt{But any fill-rank orthogonal basis would work equally well (i.e. any orthogonal transform of the PCA basis)} a PCA of $\xb_i$ with the maximum
+Let us consider $\xb_i = U_{\xb_i} \zb_i$
+%
+% \bt{But any fill-rank orthogonal basis would work equally well (i.e. any
+%   orthogonal transform of the PCA basis)}
+%
+a PCA of $\xb_i$ with the maximum
 number of components.
@@
 As the next property shows, $U_{\xb_i}$ is an optimal atlas for deterministic
-FastSRM.
+FastSRM. Note that this optimal atlas is not unique: any orthogonal matrix with
+the same span as $X_i$ would also work.
 %
@@
 
-We call probSRM($\psi$)\bt{I don't unserstand what $\psi$ represents}
-the probabilistic SRM algorithm modified such that updates
-\begin{align}
-\sigma_i^2 \leftarrow \frac1{v} (\EE[\| \xb_i - A_i \EE[\sbb|\xb]\|^2] + \| \diag(\VV[\sbb | \xb]) \|^2)
-\end{align}
-are replaced by updates
-\begin{align}
-  \sigma_i^2 \leftarrow \frac1{\psi} (\EE[\| \xb_i - A_i \EE[\sbb|\xb]\|^2] + \| \diag(\VV[\sbb | \xb]) \|^2)
-\end{align}
-
 We have the following result:
@@
   Let $(A_i)_i, \sigma_i, \Sigma_s$ be the solution obtained by probabilistic SRM on data
-  $\xb_i$ and $(A_i')_i, \sigma_i', \Sigma_s'$ the solution obtained by ProbSRM($v$) on
-  data $\zb_i = U_{\xb_i}^{\top} \xb_i$.
+  $\xb_i$ and $(A_i')_i, \sigma_i', \Sigma_s'$ the solution obtained by ProbSRM on
+  data $\zb_i = U_{\xb_i}^{\top} \xb_i$ but where updates:
+  \begin{align}
+    \sigma_i^2 \leftarrow \frac1{k} (\EE[\| \zb_i - A_i \EE[\sbb|\zb]\|^2] + \| \diag(\VV[\sbb | \zb]) \|^2)
+  \end{align}
+  are replaced by updates
+  \begin{align}
+    \sigma_i^2 \leftarrow \frac1{v} (\EE[\| \zb_i - A_i \EE[\sbb|\zb]\|^2] + \| \diag(\VV[\sbb | \zb]) \|^2)
+  \end{align}
+  .
   %
@@
 %
-We used nilearn~\cite{abraham2014machine} for fMRI data preprocessing,
-Brainiak~\cite{kumar2020brainiak2} for the non-identifiable version of SRM that
-implements the work of~\cite{chen2015reduced} and~\cite{anderson2016enabling}, Numpy~\cite{harris2020array} for
-array processing, Scipy~\cite{2020SciPy-NMeth} its implementation of the
-Hungarian algorithm, Matplotlib~\cite{hunter2007matplotlib} for plotting and
-Sklearn~\cite{pedregosa2011scikit} for machine learning pipelines.
-\bt{Please provide versions of the software used}
+We used nilearn~\cite{abraham2014machine} (version 0.8.1) for fMRI data preprocessing,
+Brainiak~\cite{kumar2020brainiak2} (version 0.9) for the non-identifiable version of SRM that
+implements the work of~\cite{chen2015reduced} and~\cite{anderson2016enabling},
+Numpy~\cite{harris2020array} (version 1.19.0) for
+array processing, Scipy~\cite{2020SciPy-NMeth} (version 1.5.3) its implementation of the
+Hungarian algorithm, Matplotlib~\cite{hunter2007matplotlib} (version 3.1.2) for plotting and
+Sklearn~\cite{pedregosa2011scikit} (version 0.23.2) for machine learning pipelines.
 
@@
 %
-The parameters $\sigma_i$, $A_i$ and $\Sigma_s$ are generated randomly.
+The parameters $\sigma_i^2$, $A_i$ and $\Sigma_s$ are generated randomly.
 %
 We sample the value of the subject specific noise level from a normal
-distribution: $\sigma_i \sim \Ncal(0, 0.1)$ \bt{$\sigma_i$ has to be non-negative}.
+distribution: $\sigma_i \sim \Ncal(0, 0.1)$
+%\bt{$\sigma_i$ has to be non-negative}.
+% No only $\sigma_i^2$.
 %
@@
 Although SHERLOCK data contains originally only 1 run, we split it into 4 blocks
-of 395 timeframes and one block of 396 timeframes for the needs of our
+of 395 time frames and one block of 396 time frames for the needs of our
 experiments.
@@
 Assuming that the data follow the model, identifiability ensures that the
-correct parameters can be recovered without ambiguity.\bt{I find this a bit ambiguous: you give te impression that you gurantee that the optimal solution is used. Rephrase.}
+problem is well posed since the parameters of the model are not ambiguous. Note however that SRM
+stays non-convex so there is no guarantee that the optimal solution is found.
+%
+% \bt{I find this a bit ambiguous: you give te impression that you gurantee that the optimal solution is used. Rephrase.}
 %
@@
 %
-We select a target time-segment (9 consecutive timeframes) in the shared
+We select a target time-segment (9 consecutive time frames) in the shared
 responses and try to localize the corresponding time segment in the components
@@
 
+\section{Implementation and reproducibility}
+Our work is fully reproducible and our code is tested and well documented. All
+material is available in our github repository \url{https://github.com/hugorichard/FastSRM}.
+
+The API is reminiscent of the scikit-learn one where the model object is
+instantiated with its parameters and then fitted on data via a method
+\emph{fit()}. After the model is fit, the learned parameters can be accessed.
+Then, a \emph{transform()} method computes the shared
+response from the learned parameters and the data passed as the argument of the \emph{transform()} method.
+
+While brainiak and nilearn are useful to reproduce some of our experiments, they
+are not strong dependencies and FastSRM only needs numpy ($\geq 1.12$), scipy
+($\geq 0.18.0$), matplotlib ($\geq 2.0.0$), scikit-learn ($geq 0.23),
+
+Our API is very simple and reminiscent of the one used in scikit-learn with a clear documentation.
+
 \section{Conclusion}
@@
 %
-This is particularly related to the development of ddep phenotyping models \bt{cite current intensive acquisition endeavours, such as courtois-Neuromod}.
+This is particularly related to the development of deep phenotyping
+models such as the Courtois project on Neural Modeling (see \url{https://docs.cneuromod.ca}).
 %

### 2021-11-25 tests/test_identifiable_srm.py
remove tqdm as requirements

@@ -13,7 +13,6 @@ from fastsrm.utils import (
     apply_aggregate,
     apply_input_format,
 )
-from tqdm import tqdm
 
 n_voxels = 300
 n_subjects = 5
@@ -341,7 +340,7 @@ def test_ncomponents():
 
 @pytest.mark.parametrize("method", ("prob", "det"))
 def test_use_pca(method):
-    for i in tqdm(range(20)):
+    for i in range(20):
         X_train = [np.random.rand(100, 10) for _ in range(3)]
         srm = IdentifiableFastSRM(
             n_components=5,

### 2021-11-25 paper/main.tex
clean requirements

@@
 are not strong dependencies and FastSRM only needs numpy ($\geq 1.12$), scipy
-($\geq 0.18.0$), matplotlib ($\geq 2.0.0$), scikit-learn ($geq 0.23),
+($\geq 0.18.0$), matplotlib ($\geq 2.0.0$), scikit-learn ($geq 0.23$), pytest
+($\geq 6.2.5$) , joblib ($\geq 1.1.0)
 

### 2021-11-25 setup.py
update FastSRM

@@ -9,22 +9,24 @@ with open("README.md", "r") as fh:
     long_description = fh.read()
 
 setup(
-    name='fastsrm',
+    name="fastsrm",
     install_requires=[
-        'scikit-learn',
-        'joblib',
-        'tqdm',
-        'matplotlib'
+        "scipy>=0.18.0"
+        "numpy>=1.12"
+        "scikit-learn>=0.23"
+        "joblib>=1.1.0"
+        "matplotlib>=2.0.0"
+        "pytest>=6.2.5"
     ],
-    version = '0.0.3',
-    license='MIT',
-    author='Hugo RICHARD',
-    download_url='https://github.com/hugorichard/FastSRM/archive/v_002.tar.gz',
-    author_email='hugo.richard@inria.fr',
-    url='https://github.com/hugorichard/FastSRM',
+    version="0.0.3",
+    license="MIT",
+    author="Hugo RICHARD",
+    download_url="https://github.com/hugorichard/FastSRM/archive/v_002.tar.gz",
+    author_email="hugo.richard@inria.fr",
+    url="https://github.com/hugorichard/FastSRM",
     long_description=long_description,
-    description='Fast Shared response model',
-    keywords='Component analysis, fMRI',
+    description="Fast Shared response model",
+    keywords="Component analysis, fMRI",
     packages=find_packages(),
-    python_requires='>=3',
+    python_requires=">=3",
 )

### 2021-11-25 paper/main.tex
update readme

--- 2021-11-25 paper/main.tex
@@
 are not strong dependencies and FastSRM only needs numpy ($\geq 1.12$), scipy
-($\geq 0.18.0$), matplotlib ($\geq 2.0.0$), scikit-learn ($geq 0.23),
-
-Our API is very simple and reminiscent of the one used in scikit-learn with a clear documentation.
+($\geq 0.18.0$), matplotlib ($\geq 2.0.0$), scikit-learn ($geq 0.23$), pytest
+($\geq 6.2.5$) , joblib ($\geq 1.1.0).
+
+The package comes with a documentation and working examples (
 

### 2021-11-25 paper/main.tex
update readme
update ppaer

@@
  it is much faster and more memory
-efficient than other implementations that do not make use of the optimal atlas.
+efficient than other implementations that do not make use of an optimal atlas.
 % on synthetic as well as on real data,
@@
 
-A complexity analysis shows that finding the optimal atlas becomes the limiting
+A complexity analysis shows that finding an optimal atlas becomes the limiting
 step of the pipeline.
@@
 of iterations has a very strong impact on performance while it has a small impact
-when the optimal atlas is used.
+when an optimal atlas is used.
 %
@@
 iterations is needed, which would yield an even higher difference in fitting time
-between methods using no atlas and methods using the optimal atlas.
+between methods using no atlas and methods using an optimal atlas.
 %
@@
 %
-When the optimal atlas is used, the accuracy is the same as when no
+When an optimal atlas is used, the accuracy is the same as when no
 atlas is used but the fitting time is reduced by a factor 10 to 100
@@
 
-The package comes with a documentation and working examples (
+The package comes with a documentation and a tutorial that reproduces, at a
+smaller scale, the experiment available in Figure~\ref{fig:srm:synthetic_gradient}.
+These are found at \url{https://hugorichard.github.io/FastSRM/index.html}.
+The instructions to fully reproduce the experiments presented in the papers are
+available in the README:
+\url{https://github.com/hugorichard/FastSRM/blob/master/README.md}.
+
+A set of tests reduces the risk of introducing a bug. These tests are run any
+time a pull request or a merged is performed on the main codebase.
 
@@
 with provably no loss of performance. 
+Note that the new memory requirements do not depend on the number of subjects.
+This is really interesting as it would make it possible to apply FastSRM on
+datasets with a very large number of subjects that do not hold in memory.
 
 We show on synthetic and real data that FastSRM is much faster and more memory
-efficient that implementations not using the optimal atlas.
+efficient that implementations not using an optimal atlas.
 Furthermore, FastSRM is identifiable. On real data, we show that compared to
@@
 %
-Our optimal atlas apprach shows that SRM-like methods do not need the
+Our optimal atlas approach shows that SRM-like methods do not need the
 full data to provide accurate estimates of shared components. We believe such
 insights may guide the design of future methods.
-\bt{Add a paragraph on implementation, sofwtare quality, documentation, examples, availability...}
-\bt{Discuss the use for other brain imaging modelities (MEEG) if relevant ?}
+%
+
+Lastly, FastSRM is useful when the number of features is much larger than the
+number of samples. This setting is common in fMRI but less common in MEG/EEG where
+the number of samples is usually much larger than the number of features. Even
+in this case, it is possible to considerably speed up SRM. Indeed, SRM only
+depends on second order statistics.
+These statistics can be pre-computed making the rest of the
+pipeline insensitive to the number of samples in the dataset. An implementation
+of this simple idea would make SRM easily applicable to other modalities.
+We live it to future work.
+% \bt{Add a paragraph on implementation, sofwtare quality, documentation, examples, availability...}
+% \bt{Discuss the use for other brain imaging modelities (MEEG) if relevant ?}
 

### 2021-11-29 paper/main.tex
update setup and paper

@@
   \begin{align}
-    \sigma_i^2 \leftarrow \frac1{k} (\EE[\| \zb_i - A_i \EE[\sbb|\zb]\|^2] + \| \diag(\VV[\sbb | \zb]) \|^2)
+    \sigma_i^2 \leftarrow \frac1{t} (\EE[\| \zb_i - A_i \EE[\sbb|\zb]\|^2] + \| \diag(\VV[\sbb | \zb]) \|^2)
   \end{align}
@@
 are not strong dependencies and FastSRM only needs numpy ($\geq 1.12$), scipy
-($\geq 0.18.0$), matplotlib ($\geq 2.0.0$), scikit-learn ($geq 0.23$), pytest
-($\geq 6.2.5$) , joblib ($\geq 1.1.0).
+($\geq 0.18.0$), matplotlib ($\geq 2.0.0$), scikit-learn ($\geq 0.23$), pytest
+($\geq 6.2.5$) , joblib ($\geq 1.1.0$).
 

### 2021-11-29 setup.py

@@ -5,8 +5,9 @@ import sys
 import setuptools
 from copy import deepcopy
 
-with open("README.md", "r") as fh:
-    long_description = fh.read()
+from pathlib import Path
+this_directory = Path(__file__).parent
+long_description = (this_directory / "README.md").read_text()
 
 setup(
     name="fastsrm",
@@ -18,13 +19,14 @@ setup(
         "matplotlib>=2.0.0"
         "pytest>=6.2.5"
     ],
-    version="0.0.3",
+    version="0.0.4",
     license="MIT",
     author="Hugo RICHARD",
-    download_url="https://github.com/hugorichard/FastSRM/archive/v_002.tar.gz",
+    download_url="https://github.com/hugorichard/FastSRM/archive/v_004.tar.gz",
     author_email="hugo.richard@inria.fr",
     url="https://github.com/hugorichard/FastSRM",
     long_description=long_description,
+    long_description_content_type='text/markdown',
     description="Fast Shared response model",
     keywords="Component analysis, fMRI",
     packages=find_packages(),

### 2021-11-30 paper/main.tex
A pass on the paper

@@
 Furthermore, SRM is
-not identifiable which makes the shared response difficult to interpret.
+not identifiable, which makes the shared response difficult to interpret.
 
@@
 We then introduce the FastSRM that relies on a dimension reduction step that we
-prove to be optimal in the sense that working with reduced data does not induce any change in the
-algorithm trajectory.
+prove to yield the same solution as the original algorithm.
+% be optimal in the sense that working with reduced data does not induce any change in the
+% algorithm trajectory.
 %
@@
 that FastSRM is considerably faster and more memory efficient
-than competitive implementations.
+than current implementations.
 
 The experiments performed in the paper are \emph{fully} reproducible: our
-code available at \url{https://github.com/hugorichard/FastSRM} downloads the
-data, runs the experiments and plot the figures.
+code available at \url{https://github.com/hugorichard/FastSRM} allows to download the
+data, run the experiments and plot the figures.
 %
@@
 model~\cite{poline2012general} cannot be used. Although some works have used
-natural annotations~\cite{huth2012continuous} or deep neural
+natural annotations~\cite{huth2012continuous}, see also \url{https://neuroscout.org} or deep neural
 networks~\cite{gucclu2017increasingly}~\cite{richard2018optimizing} to create an
@@
 %
-The gains in speed and memory usage are based on the intuition that a compressed
-representation of the input should be good enough to find a suitable estimate of
-the shared components as they live in reduced space.
-%
-It turns out that there
-exists an optimal atlas that reduces the data in a such a way that the obtained
-shared components (and other parameters) are the same as if they were computed
-on full data.
+The gains in speed and memory usage are based on the observation that a low-dimensional, yet distance-preserving
+representation of the images yields the same result as the full data.
+%
+Such a representation can be interpreted as an initial spatial  decomposition of the data.
+% that reduces the data in a such a way that the obtained
+% shared components (and other parameters) are the same as if they were computed
+% on full data.
 %
 We show that FastSRM brings several practical benefits: on real data, it produces more stable
-estimates of the shared components than its non-identifiable counter-part and 
+estimates of the shared components than its non-identifiable counterpart and 
  it is much faster and more memory
@@
 
-Our python implementation is available at
+A Python implementation is available at
 \url{https://github.com/hugorichard/FastSRM}.
@@
 \end{align}
-where $\xb_i \in \RR^v$ is the data of view $i$, $A_i \in \RR^{p \times v}$ is the
+where $\xb_i \in \RR^v$ is the data of subject $i$, called view $i$, $A_i \in \RR^{p \times v}$ is the
 mixing matrix of view $i$, $\nb_i$ is the noise of view $i$ and $\sbb \in \RR^p$ are the
@@
 %
+$p$ is the number of time frames, $v$ is the number of voxels and $m$
+is the number of subjects.
 
@@
 %
-The mixing
-matrices are spatial topographies specific to each subjects while the shared
-components give the common timecourses.
-%
-In~\cite{chen2015reduced, anderson2016enabling}, two versions of the shared
-response model are introduced which we now present.
-%
+The mixing matrices are spatial topographies specific to each subject
+while the shared components represent the common timecourses.
+%
+In~\cite{chen2015reduced, anderson2016enabling}, two versions of the
+shared response model are introduced, namely the deterministic and
+probabilistic models.
 
@@
 The likelihood is then given by: $p(\xb) = \prod_i \Ncal(\xb_i; A_i \sbb, \sigma^2 I)$ and
-therefore the empirical expected negative log-likelihood is given up to a constant independent of
+therefore the empirical expected \textbf{XXX why expected ?} negative log-likelihood is given up to a constant independent of
 $A_i$ and $S$ by:
@@
 The gradient with respect to $S$ is given by $\sum_i A_i^{\top}(A_i S -
-X_i) = \sum_i (S -
-A_i^{\top} X_i)$
+X_i) = \sum_i (S - A_i^{\top} X_i)$
 yielding the closed form updates:
@@
 \begin{equation}
-  A_i \leftarrow  \Pcal(\frac1{n}X_i S^{\top})
+  A_i \leftarrow  \Pcal \left(\frac1{n}X_i S^{\top} \right)
   \label{eq:detsrm:Aiupdate}
@@
 Indeed, after the updates with respect to each mixing matrix have been
-carried out, only the gradient with respect to $S$ remains: $\sum_i
+carried out, only the gradient with respect to $S$ remains \textbf{XXX WDYM ?}: $\sum_i
 (S - A_i^{\top}\xb_i)$.
@@
 As already highlighted in introduction, this causes the model to be
-unidentifiable (see~\cite{richard2020modeling} Appendix D for a proof). Enforcing a diagonal $\Sigma_s$ ensures identifiability (provided the diagonal values are different).
+unidentifiable (see~\cite{richard2020modeling} Appendix D for a proof). Enforcing a diagonal $\Sigma_s$ ensures identifiability, provided that the diagonal values are different.
 %
@@
 
-\subsection{Reducing the computational burden by use of atlases}
-SRM algorithms use different set of parameters $\theta$ to represent the data.
+\subsection{Reducing the computational burden by the use of spatial decompositions}
+
+\textbf{XXX Sorry but we can't call $U$ and atlas: it does not define regions, it has no sign constraints, it is not sparse etc.
+It is actually closer to a random projection matrix than two an atlas !
+So I propose to systematically remove the term atlas, and use spatial decompositions}
+
+SRM algorithms use different sets of parameters $\theta$ to represent the data.
 %
@@
 %
-The figure~\ref{fig:srm:conceptual} illustrates this process.
+Figure~\ref{fig:srm:conceptual} illustrates this process.
 %
@@
 
-\subsection{An optimal atlas}
+\subsection{A class of optimal spatial decompositions}
 \label{sec:optimal_atlas}
@@
 FastSRM. Note that this optimal atlas is not unique: any orthogonal matrix with
-the same span as $X_i$ would also work.
+the same span as the columns of $X_i$ would also work.
 %
@@
 However the
-algorithm applied on reduced data need to be slightly modified.
+algorithm applied on reduced data needs to be slightly modified.
 %
@@
 We have the following result:
-\begin{proposition}[Optimal atlas for probabilistic FastSRM]
+\begin{proposition}[Optimal spatial decomposition for probabilistic FastSRM]
   Let $(A_i)_i, \sigma_i, \Sigma_s$ be the solution obtained by probabilistic SRM on data
@@
 
-A complexity analysis shows that finding an optimal atlas becomes the limiting
+A complexity analysis shows that finding an optimal spatial decomposition becomes the limiting
 step of the pipeline.
 %
-Even with fast implementations, the subject specific PCA
+Even with fast implementations, subject specific PCA
 is costly.
@@
 $\bigO(vt^2)$ (the same as a PCA) the constant in the $\bigO$ is exactly one so
-this operation costs a lot less than the PCA
+this operation costs a lot less than a PCA
 on full data.
@@
 number of features which prevent its application to full brain data.
-In~\cite{anderson2016enabling}, the authors address this issue by using the
+In~\cite{anderson2016enabling}, this issue is addressed by using the
 inversion lemma to remove the quadratic dependence. Their implementation is now the most widely used one.
@@
 compared to the implementation in~\cite{anderson2016enabling} by the use of
-optimal atlases.
+optimal spatial decompositions.
 
@@
 %
-Algorithms are designated by the atlas they use and therefore SRM algorithms are refered to
+Algorithms are designated by the atlas they use and therefore SRM algorithms are referred to
 as \emph{None} because no atlas is used and FastSRM algorithms will have the
@@
 ~\cite{schaefer2017local, bellec2010multi, mensch2018extracting}) but
-using them does not brain any guarantee that the performance is the same as SRM.
+using them does not bring any guarantee that the performance is the same as SRM.
 %
@@
 Assuming that the data follow the model, identifiability ensures that the
-problem is well posed since the parameters of the model are not ambiguous. Note however that SRM
-stays non-convex so there is no guarantee that the optimal solution is found.
+problem is well posed since the solutions are well-characterized. Note however that SRM
+remains non-convex so there is no guarantee that the optimal solution is found.
 %
@@
 model is used.
-%
-
-To measure the stability of the common components obtained from the Sherlock
+
+To measure the stability of the shared components obtained from the Sherlock
 dataset, we divide the subjects into two roughly equal groups and extract the
@@
 The procedure is repeated 9 times using the Brainiak implementation (that is not
-identifiable) and our
+identifiable since the shared components covariance is unconstrained) and our
 FastSRM implementation.
@@
 We plot the results as a scatter plot in Figure~\ref{exp:identifiability}, where, for each repetition , the x-axis indicates the stability measure obtained with Brainiak's implementation
-(not identifiable since the shared components covariance is unconstrained) and the y-axis the stability measure
+ and the y-axis the stability measure
 obtained with FastSRM (identifiable since the shared components covariance is diagonal).
@@
 Furthermore, FastSRM is identifiable. On real data, we show that compared to
-Brainiak's implementation (which is not identifiable), FastSRM provides more
+Brainiak's implementation, that is not identifiable, FastSRM provides more
 stable estimates of the shared components.
@@
 FastSRM inherits from all the weaknesses of SRM including the fact that
-mixing matrices are assumed to be orthogonal which is rather unrealistic. It
+mixing matrices are assumed to be orthogonal, which is rather unrealistic. It
 remains to be seen whether dimension reduction and learning of the shared
@@
 of this simple idea would make SRM easily applicable to other modalities.
-We live it to future work.
+We leave it to future work.
 % \bt{Add a paragraph on implementation, sofwtare quality, documentation, examples, availability...}

### 2021-12-01 paper/main.tex
adress comments

@@
 The likelihood is then given by: $p(\xb) = \prod_i \Ncal(\xb_i; A_i \sbb, \sigma^2 I)$ and
-therefore the empirical expected \textbf{XXX why expected ?} negative log-likelihood is given up to a constant independent of
+therefore the empirical
+%
+% \textbf{XXX why expected ?} It was for empirical expectation but it may be
+% confusing so I removed it.
+%
+negative log-likelihood is given up to a constant independent of
 $A_i$ and $S$ by:
@@
 Indeed, after the updates with respect to each mixing matrix have been
-carried out, only the gradient with respect to $S$ remains \textbf{XXX WDYM ?}: $\sum_i
-(S - A_i^{\top}\xb_i)$.
+carried out, the gradient with respect to $(W_i)_{i=1}^m$ is zero and therefore
+to compute the $\ell_{\infty}$ norm of the gradient we only need the gradient
+with respect to $S$: $m S - \sum_i A_i^{\top}\xb_i$ where the right hand side is
+used in the updates of $S$.
 %

### 2021-12-13 paper/main.tex
remove reference to atlases

@@
 \usepackage{amsmath}
+\usepackage{thmtools, thm-restate}
 \usepackage{utils/kky}
@@
 ICA assumes that components are independent, a rather realistic assumption which
-ensures identifiability up to permutation and scaling of the model and can be fit efficiently.
+ensures identifiability up to permutation and scaling of the model and can be fitted efficiently.
 %
@@
 In this work, we introduce FastSRM, an identifiable model that uses an optimal
-atlas-based decomposition to speed up the computations with provably no loss of
+spatial decomposition to speed up the computations with provably no loss of
 performance.
@@
  it is much faster and more memory
-efficient than other implementations that do not make use of an optimal atlas.
+efficient than other implementations that do not make use of an optimal spatial decomposition.
 % on synthetic as well as on real data,
@@
 
-\textbf{XXX Sorry but we can't call $U$ and atlas: it does not define regions, it has no sign constraints, it is not sparse etc.
+% \textbf{XXX Sorry but we can't call $U$ and atlas: it does not define regions, it has no sign constraints, it is not sparse etc.
 It is actually closer to a random projection matrix than two an atlas !
@@
 
-In fMRI, the classical approach used to reduce the data is to apply an atlas.
-%
-A deterministic atlas such as~\cite{bellec2010multi} is a parcellation of the
+In fMRI, spatial decompositions are often used to reduce the dimensionality of
+the data.
+%
+A classical approach is to apply a deterministic atlas such
+as~\cite{bellec2010multi} which is a parcellation of the
 brain into $r$ regions.
 %
-Reducing an image using a deterministic atlas corresponds to
-averaging the signal within each region of the atlas.
-%
-A probabilistic atlases such
-as~\cite{dadi_fine-grain_2020} describes each region as a set of weights across
+There also exists probabilistic atlases
+such as~\cite{dadi_fine-grain_2020} that describes each region as a set of weights across
 the full brain. 
 %
-Therefore, the image reduction can be done with a matrix product.
-%
-
-In FastSRM we consider a set of view-specific atlases $U_i \in \RR^{v \times r}$ such that
-$U_i^{\top}U_i = I_r$ where $r$ is the number of regions in the atlas.
+
+Deterministic and probabilistic atlases are spatial decompositions that do not
+depend on the view at hand. In FastSRM, we consider a set of view-specific
+spatial decompositions
+$U_i \in \RR^{v \times r}$ such that
+$U_i^{\top}U_i = I_r$ where $r$ is the number of components in the spatial decompositions.
 %
@@
 %
-However, as we will see in the next section, there exists atlases such that the
+However, as we will see in the next section, there exists spatial decompositions such that the
 models parametrized by $\theta$ and
@@
   \includegraphics[scale=0.34]{figures/srm/conceptual_figure2.pdf}
-  \caption{\textbf{FastSRM algorithm} In step 1, data $\xb_i$ are projected onto an
-    atlas $U_i$ that may depend on the subject $i$ (top).
+  \caption{\textbf{FastSRM algorithm} In step 1, data $\xb_i$ are projected onto a
+    spatial decomposition $U_i$ that may depend on the subject $i$ (top).
     %
@@
 once, it is no longer necessary to keep the full dataset in memory (we can load
-data $X_i$ one after the other and similarly for the atlases $U_i$).
+data $X_i$ one after the other and similarly for the spatial decomposition $U_i$).
 %
@@
 %
-The number of regions in large atlases is about $r=1000$ and in full brain data,
+The number of components in large spatial decompositions is about $r=1000$ and in full brain data,
 the number of voxels is about $300~000$ so that $\frac{v}{r}$ is typically about
@@
 \label{sec:optimal_atlas}
-In principle, FastSRM can be used with any atlas.
+In principle, FastSRM can be used with any spatial decomposition.
 %
 However, in general, working with reduced data induces a loss of information
-that can be minimized if the atlas is carefully chosen.
+that can be minimized if the spatial decomposition is carefully chosen.
 %
@@
 %
-Yet, we show that there exists an optimal atlas in the
+Yet, we show that there exists an optimal spatial decomposition in the
 sense that SRM and FastSRM yield the same results.
@@
 %
-Therefore $U_{\xb_i}$ constitutes a possible choice of subject specific atlas.
-%
-As the next property shows, $U_{\xb_i}$ is an optimal atlas for deterministic
-FastSRM. Note that this optimal atlas is not unique: any orthogonal matrix with
+Therefore $U_{\xb_i}$ constitutes a possible choice of subject specific spatial decomposition.
+%
+As the next property shows, $U_{\xb_i}$ is an optimal spatial decomposition for deterministic
+FastSRM. Note that this optimal spatial decomposition is not unique: any orthogonal matrix with
 the same span as the columns of $X_i$ would also work.
@@
 
-\begin{proposition}[Optimal atlas for deterministic FastSRM]
+\begin{proposition}[Optimal spatial decomposition for deterministic FastSRM]
   Let $(A_i)_i, S$ be the solution obtained by deterministic SRM on data
   $(X_i)_i$ and $(A_i')_i, S'$ the solution obtained by deterministic FastSRM on
-  data $(X_i)_i$ using atlases $(U_{X_i})_i$ where $X_i = U_{\xb_i} Z_i$ is a PCA
+  data $(X_i)_i$ using spatial decompositions $(U_{X_i})_i$ where $X_i = U_{\xb_i} Z_i$ is a PCA
   of $X_i$.
@@
 \end{proposition}
-\begin{proof}
-Updates of the mixing matrices $A_i$ in deterministic SRM
-equation~\eqref{eq:detsrm:Aiupdate} can be written:
-\begin{align}
-  A_i &\leftarrow \Pcal \left(\frac1{n}X_i S^{\top} \right)= U_{X_i}\Pcal \left(\frac1{n}Z_i S^{\top} \right)
-  \label{eq:fastdetsrm:Aiupdate}
-\end{align}
-where $\Pcal$ is the projection on the Stiefel manifold: $\Pcal(M) = M
-(M^{\top}M)^{-\frac12}$.
-%
-
-Therefore we can look for $A_i$ as $A_i = U_{X_i} \tilde{A}_i$.
-%
-We can show that $\tilde{A}_i$ is orthogonal.
-%
-Indeed,
-\begin{align}
-  &A_i^{\top} A_i = I_p \\
-  & \implies \tilde{A}_i^{\top}U_{X_i}^{\top} U_{X_i} \tilde{A}_i = I_p \\
-  & \implies \tilde{A}_i^{\top} \tilde{A}_i = I_p
-\end{align}
-
-Then, we use the fact that
-\begin{align}
-  \|X_i - A_i S \|^2 = \| U_{X_i}Z_i - U_{X_i}\tilde{A}_i S\|^2 = \| Z_i - \tilde{A}_i S \|^2
-  \label{eq:equality:xy}
-\end{align}
-so that $A_i' = \tilde{A}_i$.
-%
-
-Therefore, the solution of deterministic SRM on data $(\zb_i)_{i=1}^m$ and
-$(\xb_i)_{i=1}^m$ are linked by the change of parameters $A_i = U_{\xb_i}A_i'$ and
-$S = S'$.
-%
-This concludes the proof.
-%
-
-\end{proof}
-
+The proof is available in appendix~\ref{proof:deterministic}.
 In the case of probabilistic SRM we can obtain very similar results.
@@
   \end{align}
-  .
   %
@@
 \end{proposition}
-\begin{proof}
-  Updates of the mixing matrices $A_i$ in probabilistic SRM
-  equation~\eqref{eq:psrm:Aiupdate} can be written:
-  \begin{align}
-    A_i &\leftarrow U_{\xb_i}\Pcal(\EE[\zb_i \EE[\sbb| \xb_i]^T])
-    \label{eq:fastprobsrm:Aiupdate}
-  \end{align}
-  so we can look for $A_i$ as $A_i = U_{\xb_i} \tilde{A}_i$ and, as in the
-  deterministic case, $\tilde{A}_i$ is orthogonal.
-  %
-  Therefore equality~\eqref{eq:equality:xy} holds.
-  %
-
-  Then we consider the expected negative log-likelihood of probabilistic srm:
-  \begin{align}
-    \loss &= \sum_i \frac12 v\log(\sigma_i^2) + \frac12 \log(|\Sigma_s|) + \EE[\int_{\sbb} \sum_i \frac1{2 \sigma_i^{2}} \|\xb_i - A_i \sbb \|^2 \nonumber \\& \enspace \enspace \enspace \enspace+ \frac12 \dotp{ \sbb }{ \Sigma_s^{-1} \sbb }  d\sbb] \\
-          &= \sum_i \frac12 v \log(\sigma_i^2) + \frac12 \log(|\Sigma_s|) + \EE[\int_{\sbb} \sum_i \frac1{2 \sigma_i^{2}}\|\zb_i - \tilde{A}_i \sbb \|^2 \nonumber \\& \enspace \enspace \enspace \enspace+ \frac12 \dotp{ \sbb }{ \Sigma_s^{-1} \sbb }  d\sbb]
-  \end{align}
-  where we use equality~\eqref{eq:equality:xy}.
-  %
-
-  Optimizing the log-likelihood via expectation maximization yields the exact
-  same updates as probabilistic SRM on data $\zb_i$
-  except that updates
-  \begin{align}
-    \sigma_i^2 \leftarrow \frac1{t} (\EE[\| \zb_i - \tilde{A}_i \EE[\sbb|\zb]\|^2] + \| \diag(\VV[\sbb | \zb]) \|^2)
-  \end{align}
-  are replaced by updates
-  \begin{align}
-    \sigma_i^2 \leftarrow \frac1{v} (\EE[\| \zb_i - \tilde{A}_i \EE[\sbb|\zb]\|^2] + \| \diag(\VV[\sbb | \zb]) \|^2)
-  \end{align}
-  so that $\tilde{A}_i = A'_i$.
-  %
-
-  Therefore, the updates in both algorithms are linked  by $A_i = U_{\xb_i}A_i'$,
-  $\sigma_i' =
-  \sigma_i$ and $\Sigma_s'  = \Sigma_s$.
-  %
-
-  This concludes the proof.
-  %
-
-\end{proof}
-
+  The proof is available in appendix~\ref{proof:prob}.
 Proposition~\ref{prop:optimaldetsrm} and Proposition~\ref{prop:optimalprobsrm} show that
@@
 %
-Algorithms are designated by the atlas they use and therefore SRM algorithms are referred to
-as \emph{None} because no atlas is used and FastSRM algorithms will have the
+Algorithms are designated by the spatial decomposition they use and therefore SRM algorithms are referred to
+as \emph{None} because no spatial decomposition is used and FastSRM algorithms will have the
 label \emph{Optimal}.
@@
 Note that it would be possible to use FastSRM with sub-optimal
-atlases (there exists a wide variety of brain atlases
+spatial decompositions (there exists a wide variety of brain spatial decompositions
 ~\cite{schaefer2017local, bellec2010multi, mensch2018extracting}) but
@@
 
-We empirically see that the optimal approach is equivalent to using no atlas in
+We empirically see that the optimal approach is equivalent to using no spatial decompositions in
 terms of performance.
@@
 %
-When no atlas is used, the number
+When no spatial decompositions is used, the number
 of iterations has a very strong impact on performance while it has a small impact
-when an optimal atlas is used.
+when an optimal spatial decomposition is used.
 %
@@
 iterations is needed, which would yield an even higher difference in fitting time
-between methods using no atlas and methods using an optimal atlas.
+between methods using no spatial decomposition and methods using an optimal
+spatial decomposition.
 %
@@
     %
-    As expected, when optimal atlases are used,
-    the performance is the same as if no atlas is used but the fitting time is
+    As expected, when optimal spatial decompositions are used,
+    the performance is the same as if no spatial decomposition is used but the fitting time is
     much lower.
@@
 %
-Then we show that our FastSRM algorithm (that works with optimal atlases)
+Then we show that our FastSRM algorithm (that works with optimal spatial decompositions)
 matches the performance of SRM (that works on full data) but use much less
@@
 %
-When an optimal atlas is used, the accuracy is the same as when no
-atlas is used but the fitting time is reduced by a factor 10 to 100
+When an optimal spatial decomposition is used, the accuracy is the same as when no
+spatial decomposition is used but the fitting time is reduced by a factor 10 to 100
 and so is the memory usage (see Figure~\ref{fig:srm:timesegment},
@@
     matching accuracy (middle) Fitting time (bottom) Memory usage. When the
-    optimal atlas is used, the accuracy is the same as when no atlas is used but
+    optimal spatial decomposition is used, the accuracy is the same as when no
+    spatial decomposition is used but
     the fitting time is reduced by a factor 10 to 100 and so is the memory usage.
@@
 FastSRM is an efficient implementation of SRM
-that uses optimal atlases to speed up computations and reduce memory requirements
+that uses optimal spatial decompositions to speed up computations and reduce memory requirements
 with provably no loss of performance. 
@@
 We show on synthetic and real data that FastSRM is much faster and more memory
-efficient that implementations not using an optimal atlas.
+efficient that implementations not using an optimal spatial decomposition.
 Furthermore, FastSRM is identifiable. On real data, we show that compared to
@@
 %
-Our optimal atlas approach shows that SRM-like methods do not need the
+Our approach based on optimal spatial decomposition shows that SRM-like methods do not need the
 full data to provide accurate estimates of shared components. We believe such
@@
 \bibliography{utils/biblio}
+
+\newpage
+\appendix
+
+\section{Proofs of Proposition~\ref{prop:optimaldetsrm}}
+\label{proof:deterministic}
+
+\begin{proof}
+Updates of the mixing matrices $A_i$ in deterministic SRM
+equation~\eqref{eq:detsrm:Aiupdate} can be written:
+\begin{align}
+  A_i &\leftarrow \Pcal \left(\frac1{n}X_i S^{\top} \right)= U_{X_i}\Pcal \left(\frac1{n}Z_i S^{\top} \right)
+  \label{eq:fastdetsrm:Aiupdate}
+\end{align}
+where $\Pcal$ is the projection on the Stiefel manifold: $\Pcal(M) = M
+(M^{\top}M)^{-\frac12}$.
+%
+
+Therefore we can look for $A_i$ as $A_i = U_{X_i} \tilde{A}_i$.
+%
+We can show that $\tilde{A}_i$ is orthogonal.
+%
+Indeed,
+\begin{align}
+  &A_i^{\top} A_i = I_p \\
+  & \implies \tilde{A}_i^{\top}U_{X_i}^{\top} U_{X_i} \tilde{A}_i = I_p \\
+  & \implies \tilde{A}_i^{\top} \tilde{A}_i = I_p
+\end{align}
+
+Then, we use the fact that
+\begin{align}
+  \|X_i - A_i S \|^2 = \| U_{X_i}Z_i - U_{X_i}\tilde{A}_i S\|^2 = \| Z_i - \tilde{A}_i S \|^2
+  \label{eq:equality:xy}
+\end{align}
+so that $A_i' = \tilde{A}_i$.
+%
+
+Therefore, the solution of deterministic SRM on data $(\zb_i)_{i=1}^m$ and
+$(\xb_i)_{i=1}^m$ are linked by the change of parameters $A_i = U_{\xb_i}A_i'$ and
+$S = S'$.
+%
+This concludes the proof.
+%
+
+\end{proof}
+
+\section{Proofs of Proposition~\ref{prop:optimalprobsrm}}
+\label{proof:prob}
+\begin{proof}
+  Updates of the mixing matrices $A_i$ in probabilistic SRM
+  equation~\eqref{eq:psrm:Aiupdate} can be written:
+  \begin{align}
+    A_i &\leftarrow U_{\xb_i}\Pcal(\EE[\zb_i \EE[\sbb| \xb_i]^T])
+    \label{eq:fastprobsrm:Aiupdate}
+  \end{align}
+  so we can look for $A_i$ as $A_i = U_{\xb_i} \tilde{A}_i$ and, as in the
+  deterministic case, $\tilde{A}_i$ is orthogonal.
+  %
+  Therefore equality~\eqref{eq:equality:xy} holds.
+  %
+
+  Then we consider the expected negative log-likelihood of probabilistic srm:
+  \begin{align}
+    \loss &= \sum_i \frac12 v\log(\sigma_i^2) + \frac12 \log(|\Sigma_s|) + \EE[\int_{\sbb} \sum_i \frac1{2 \sigma_i^{2}} \|\xb_i - A_i \sbb \|^2 \nonumber \\& \enspace \enspace \enspace \enspace+ \frac12 \dotp{ \sbb }{ \Sigma_s^{-1} \sbb }  d\sbb] \\
+          &= \sum_i \frac12 v \log(\sigma_i^2) + \frac12 \log(|\Sigma_s|) + \EE[\int_{\sbb} \sum_i \frac1{2 \sigma_i^{2}}\|\zb_i - \tilde{A}_i \sbb \|^2 \nonumber \\& \enspace \enspace \enspace \enspace+ \frac12 \dotp{ \sbb }{ \Sigma_s^{-1} \sbb }  d\sbb]
+  \end{align}
+  where we use equality~\eqref{eq:equality:xy}.
+  %
+
+  Optimizing the log-likelihood via expectation maximization yields the exact
+  same updates as probabilistic SRM on data $\zb_i$
+  except that updates
+  \begin{align}
+    \sigma_i^2 \leftarrow \frac1{t} (\EE[\| \zb_i - \tilde{A}_i \EE[\sbb|\zb]\|^2] + \| \diag(\VV[\sbb | \zb]) \|^2)
+  \end{align}
+  are replaced by updates
+  \begin{align}
+    \sigma_i^2 \leftarrow \frac1{v} (\EE[\| \zb_i - \tilde{A}_i \EE[\sbb|\zb]\|^2] + \| \diag(\VV[\sbb | \zb]) \|^2)
+  \end{align}
+  so that $\tilde{A}_i = A'_i$.
+  %
+
+  Therefore, the updates in both algorithms are linked  by $A_i = U_{\xb_i}A_i'$,
+  $\sigma_i' =
+  \sigma_i$ and $\Sigma_s'  = \Sigma_s$.
+  %
+
+  This concludes the proof.
+  %
+
+\end{proof}
 \end{document}

### 2021-12-15 paper/main.tex
small pass

@@
 %
-We then introduce the FastSRM that relies on a dimension reduction step that we
+We then introduce FastSRM, that relies on a dimension reduction step that we
 prove to yield the same solution as the original algorithm.
@@
 paradigms are unconstrained from behavioral manipulations and thus, more
-ecological with respect to real-world conditions.
-%
-However and in contrast to more traditional designs, these so-called
+ecological with respect to every-day cognitive conditions.
+%
+Yet, in contrast to standard designs, these so-called
 naturalistic stimuli do not come with a design matrix that quantifies the
@@
 Without a design matrix, models such as the general linear
-model~\cite{poline2012general} cannot be used. Although some works have used
-natural annotations~\cite{huth2012continuous}, see also \url{https://neuroscout.org} or deep neural
-networks~\cite{gucclu2017increasingly}~\cite{richard2018optimizing} to create an
-estimate of the design matrix of naturalistic
-stimuli it is by no means a simple task.
+model~\cite{poline2012general} cannot be used.
+%
+Although some works have used manual
+annotations~\cite{huth2012continuous}, see also
+\url{https://neuroscout.org} or deep neural
+networks~\cite{gucclu2017increasingly}~\cite{richard2018optimizing} to
+create an estimate of the design matrix of naturalistic stimuli it is
+a hard task.
 
@@
 %
-In the case where the dataset contains only one subject
+In the case where the dataset contains only one subject,
 independent component analysis~\cite{jutten1991blind} (ICA) is the method of choice.
 %
-ICA assumes that components are independent, a rather realistic assumption which
+ICA assumes that components are independent, a defendible assumption which
 ensures identifiability up to permutation and scaling of the model and can be fitted efficiently.
@@
 Many different methods
-can produce shared components from different subjects. Some assume independent
-components~\cite{richard2021model}~\cite{richard2020modeling}~\cite{varoquaux2009canica}~\cite{calhoun2001method}.
-while others are only based on second order information such as Multiset
-Canonical Correlation Analysis~\cite{via2011joint} or the Shared Response Model (SRM)~\cite{chen2015reduced}.
-%
-
-This work focuses on the SRM. Because of its
-simplicity and its built-in dimension reduction, this model is widely
-used~\cite{baldassano2017discovering}~\cite{cohen2017computational}~\cite{baldassano2018representation}~\cite{jolly2020custom}~\cite{lee2021anticipation}
-including as a preprocessing step for source
+can produce shared components from different subjects.
+%
+Some assume independent
+components~\cite{richard2021model}~\cite{richard2020modeling}~\cite{varoquaux2009canica}~\cite{calhoun2001method},
+while others are only based on second order information such as
+Multiset Canonical Correlation Analysis~\cite{via2011joint} or the
+Shared Response Model (SRM)~\cite{chen2015reduced}.
+%
+
+This work focuses on the SRM. Because of its simplicity and its
+built-in dimension reduction, this model is widely
+used~\cite{baldassano2017discovering}~\cite{cohen2017computational}~\cite{baldassano2018representation}~\cite{jolly2020custom}~\cite{lee2021anticipation},
+in particular, as a preprocessing step for source
 identification~\cite{richard2021model}~\cite{richard2020modeling},
-classification~\cite{turek2018capturing}~\cite{chen2017shared}\cite{zhang2016searchlight} or as a basis for transfer learning~\cite{zhang2018transfer}.
+classification~\cite{turek2018capturing}~\cite{chen2017shared}\cite{zhang2016searchlight}
+or as a basis for transfer learning~\cite{zhang2018transfer}.
 
@@
 We show that FastSRM brings several practical benefits: on real data, it produces more stable
-estimates of the shared components than its non-identifiable counterpart and 
+estimates of the shared components than its non-identifiable counterpart, and 
  it is much faster and more memory
@@
 \label{sec:deterministicsrm}
-Let us consider $n$ observations of $\xb_i$ and $\sbb$ that we stack into
+Let us consider $n$ observations (scans) of $\xb_i$ and $\sbb$ that we stack into
 matrices $X_i \in \RR^{v, n}$ and $S \in \RR^{p, n}$.
-
+%
 The deterministic shared response model considers both the mixing matrices $A_i$ and
@@
 
-The complexity of Deterministic SRM is in $\bigO(\mathrm{n_{iter}} mpvn)$ where
-$n$ is the number of samples and $\mathrm{n_{iter}}$ the number of iterations.
-%
-We monitor the convergence by looking at the $\ell_{\infty}$ norm of the
+The complexity of Deterministic SRM is in $\bigO(\mathrm{n_{iter}} mpvn)$, where
+ $\mathrm{n_{iter}}$ is the number of iterations.
+%
+We monitor the convergence by computing the $\ell_{\infty}$ norm of the
 gradient.
@@
 \begin{align}
-  c_2(\xb) = c_1 \exp \left( -\frac12( \sum_i
-  \frac1{\sigma_i^2}\|\xb_i\|^2 - \dotp{  \EE[\sbb }{ \xb], \VV[\sbb | \xb]^{-1}
-  \EE[\sbb | \xb] }) \right)
+  c_2(\xb) = c_1 \exp \left( -\frac12 \left( \sum_i \frac1{\sigma_i^2}\|\xb_i\|^2 - \dotp{  \EE[\sbb }{ \xb], \VV[\sbb | \xb]^{-1} \EE[\sbb | \xb] } \right) \right)
 \end{align}
@@
 %
-It remains to be shown how to draw a correspondence between FastSRM and SRM which is
+It remains to be shown how to draw a correspondence between FastSRM and SRM, which is
 addressed in the following section.
@@
 %
-In practice the cost of the matrix products $X_i^{\top} X_i$ is often still the
+In practice the cost of the matrix products $X_i^{\top} X_i$ is often the
 limiting step of the pipeline (this depends on the number of iterations) but as
@@
 %
-Note that if memory allows it, these matrix products can be
+Note that if large  memory resources are available, these matrix products can be
 computed in parallel.
@@
 The implementation of SRM proposed in~\cite{chen2015reduced} is quadratic in the
-number of features which prevent its application to full brain data.
+number of features, which prevents its application to full brain data.
 In~\cite{anderson2016enabling}, this issue is addressed by using the
@@
 
-We see SRM as a well principled formulation of the hyperalignment
+We consider SRM as a well principled formulation of the hyperalignment
 method~\cite{haxby2011common}.
@@
 In this work, we do not discuss the merit of FastSRM with respect to
-other methods. We only see that SRM is widely used and aim to provide
+other methods. We simply observe that SRM is widely used and aim to provide
 a faster, more memory efficient and identifiable implementation.  In
-our view FastSRM should be seen as a plugin replacement for SRM.
+our view FastSRM, should be used as a plugin replacement for SRM.
 
@@
 %
-We used nilearn~\cite{abraham2014machine} (version 0.8.1) for fMRI data preprocessing,
+We used Nilearn~\cite{abraham2014machine} (version 0.8.1) for fMRI data preprocessing,
 Brainiak~\cite{kumar2020brainiak2} (version 0.9) for the non-identifiable version of SRM that
@@
 In real fMRI data, the model cannot be expected to hold perfectly, but we can
-hope for greater stability in the parameters recovered than if an unidentifiable
+hope for greater stability in the parameters recovered than if an non-identifiable
 model is used.
@@
 While brainiak and nilearn are useful to reproduce some of our experiments, they
-are not strong dependencies and FastSRM only needs numpy ($\geq 1.12$), scipy
-($\geq 0.18.0$), matplotlib ($\geq 2.0.0$), scikit-learn ($\geq 0.23$), pytest
-($\geq 6.2.5$) , joblib ($\geq 1.1.0$).
+are not strong dependencies and FastSRM only needs Numpy ($\geq 1.12$), Scipy
+($\geq 0.18.0$), Matplotlib ($\geq 2.0.0$), Scikit-learn ($\geq 0.23$), Pytest
+($\geq 6.2.5$) , and Joblib ($\geq 1.1.0$).
 
@@
 %
-This is what FastSRM provides.
+This is precisely what FastSRM provides.
 
@@
 FastSRM inherits from all the weaknesses of SRM including the fact that
-mixing matrices are assumed to be orthogonal, which is rather unrealistic. It
+mixing matrices are assumed to be orthogonal, which is debattable in practice. It
 remains to be seen whether dimension reduction and learning of the shared
@@
   %
-
+  %
   Optimizing the log-likelihood via expectation maximization yields the exact
@@
   %
-
   Therefore, the updates in both algorithms are linked  by $A_i = U_{\xb_i}A_i'$,
@@
   %
-
   This concludes the proof.
   %
-
 \end{proof}

### 2022-01-04 paper/main.tex
pass on the paper with some reorganization to lighten main text

@@
 Although some works have used manual
-annotations~\cite{huth2012continuous}, see also
-\url{https://neuroscout.org} or deep neural
-networks~\cite{gucclu2017increasingly}~\cite{richard2018optimizing} to
-create an estimate of the design matrix of naturalistic stimuli it is
-a hard task.
+annotations~\cite{huth2012continuous} or deep neural
+networks~\cite{gucclu2017increasingly}~\cite{richard2018optimizing},
+see also \url{https://neuroscout.org}, to create an estimate of the
+design matrix of naturalistic stimuli, it is a hard task.
 
@@
 %
-ICA assumes that components are independent, a defendible assumption which
-ensures identifiability up to permutation and scaling of the model and can be fitted efficiently.
-%
-
-But when the dataset contains multiple subjects, a natural assumption is to
-assume that the design matrix is shared across subjects.
-%
-Many different methods
-can produce shared components from different subjects.
+ICA assumes that components are independent, a defendible assumption
+which ensures identifiability up to permutation and scaling of the
+model and can be fitted efficiently.
+%
+When the dataset contains multiple subjects undergoing the same
+protocol, a natural assumption is to assume that the design matrix is
+shared across subjects.
+%
+Many different methods can produce shared components from different
+subjects.
 %
@@
 %
+We give all details on the formulation and derivation of update rules in section \ref{details:probsrm}.
+%
 Denoting $\VV[\sbb | \xb] = (\sum_i \frac1{\sigma_i^2} I +
 \Sigma_s^{-1})^{-1}$ and $\EE[\sbb | \xb] = \VV[\sbb | \xb] \sum_i \frac1{\sigma_i^2}
-A_i^{\top}\xb_i$, we have
-\begin{align}
-  p(\xb, \sbb) &= \prod_i \frac{\exp \left(-\frac{\|\xb_i - A_i \sbb \|^2}{2 \sigma_i^2} \right) }{(2 \pi \sigma_i^{2v})^{\frac12}} \frac{\exp(-\frac12 \dotp{ \sbb }{ \Sigma_s^{-1} \sbb } )}{(2 \pi | \Sigma_s|)^{\frac12}} \\
-               &= c_1 \exp (-\frac12 \left( \sum_i \frac1{\sigma_i^2}\|\xb_i\|^2 - 2  \dotp{ \sum_i \frac1{\sigma_i^2} A_i^{\top}\xb_i}{ \sbb } \right.
-\\& \left.+ \sum_i \frac1{\sigma_i^2} \| \sbb \|^2 + \dotp{ \sbb}{ \Sigma_s^{-1} \sbb }  \right)) \\
-               &= c_2(\xb) \exp(-\frac12 \left( \dotp{  \sbb - \EE[\sbb }{ \xb], \VV[\sbb | \xb]^{-1} ( \sbb - \EE[\sbb | \xb])  } \right)) \label{probsrmcompleted}
-\end{align}
-where
-\begin{align}
-  c_1 = \frac1{(2 \pi \sigma_i^{2v})^{\frac12}}\frac1{(2 \pi |
-  \Sigma_s|)^{\frac12}}
-\end{align}
-and
-\begin{align}
-  c_2(\xb) = c_1 \exp \left( -\frac12 \left( \sum_i \frac1{\sigma_i^2}\|\xb_i\|^2 - \dotp{  \EE[\sbb }{ \xb], \VV[\sbb | \xb]^{-1} \EE[\sbb | \xb] } \right) \right)
-\end{align}
-are independent from $\sbb$.
-%
-Therefore,
-\begin{align}
-  \sbb| \xb \sim \Ncal(\EE[\sbb | \xb], \VV[\sbb, \xb])
-\end{align}
-
-The negative expected completed log-likelihood is given by
-\begin{align}
-	\loss = \sum_i \frac12 v \log(\sigma_i^2) + \frac1{2 \sigma_i^2} \EE[\| \xb_i - A_i \sbb \|^2]
-\end{align}
-updates are therefore given by:
+A_i^{\top}\xb_i$, the updates are given by:
 \begin{align}
@@
 
-It is useful to access the log-likelihood to check the implementation of the
-algorithm and monitor the convergence.
-%
-From equation~\eqref{probsrmcompleted},
-the likelihood is given by:
-\begin{align}
-  p(\xb) &= c_2(\xb) \int_{\sbb} \exp(-\frac12 \left( \dotp{  \sbb - \EE[\sbb }{ \xb], \VV[\sbb | \xb]^{-1} ( \sbb - \EE[\sbb | \xb])  } \right)) d\sbb \\
-         &= c_2(\xb) (2 \pi |\VV[\sbb | \xb]|)^{\frac12}
-\end{align}
-replacing $c_2(\xb)$ by its expression and taking the log, the expected negative
-log-likelihood is (up to constants) given by:
-\begin{align}
-  \EE[-\log(&p(\xb))] = \sum_i \frac{v}{2} \log(\sigma_i^2) + \frac12 \log(|\Sigma_s|) - \frac12 \log(|\VV[\sbb | \xb]|) \nonumber \\ &+ \sum_i
-  \frac12 \frac1{\sigma_i^2} \EE[\|\xb_i\|^2] - \frac12 \EE[\dotp{  \EE[\sbb }{ \xb], \VV[\sbb | \xb]^{-1} \EE[\sbb | \xb] }]
-\end{align}
-
 The complexity of Probabilistic SRM is $\bigO(\mathrm{n_{iter}} mpvn)$, the same as in
@@
 %
-We can monitor the convergence by looking at the log-likelihood decrease at each iteration
+We can monitor the convergence by computing the log-likelihood decrease at each iteration
 and stop the algorithm when the magnitude of the decrease is below some
@@
 %
-There also exists probabilistic atlases
+There also exist probabilistic atlases
 such as~\cite{dadi_fine-grain_2020} that describes each region as a set of weights across
@@
 %
-However, as we will see in the next section, there exists spatial decompositions such that the
+However, as we will see in the next section, there exist spatial decompositions such that the
 models parametrized by $\theta$ and
@@
 %
-It remains to be shown how to draw a correspondence between FastSRM and SRM, which is
-addressed in the following section.
+It remains to be shown how to draw a correspondence between FastSRM
+and SRM, which we do in the following section.
 %
@@
 %
-
 We have the following result:
@@
 a faster, more memory efficient and identifiable implementation.  In
-our view FastSRM, should be used as a plugin replacement for SRM.
+our view, FastSRM should be used as a plugin replacement for SRM.
 
@@
 %
-This is predicted by the theory in section~\ref{sec:optimal_atlas} where we demonstrate that these two algorithms yield exactly the same output
-from the same input.
+This is predicted by the theory in section~\ref{sec:optimal_atlas},
+where we demonstrate that these two algorithms yield exactly the same
+output from a given input and initialization.
 %
@@
 %
-When no spatial decompositions is used, the number
+When no spatial decomposition is used, the number
 of iterations has a very strong impact on performance while it has a small impact
@@
 %
-% 
-Although SHERLOCK data contains originally only 1 run, we split it into 4 blocks
+Although SHERLOCK originally data contain 1 run only, we split them into 4 blocks
 of 395 time frames and one block of 396 time frames for the needs of our
@@
 Assuming that the data follow the model, identifiability ensures that the
-problem is well posed since the solutions are well-characterized. Note however that SRM
+problem is well-posed since the solutions are well-characterized. Note however that SRM
 remains non-convex so there is no guarantee that the optimal solution is found.
@@
 %
-We plot the results as a scatter plot in Figure~\ref{exp:identifiability}, where, for each repetition , the x-axis indicates the stability measure obtained with Brainiak's implementation
- and the y-axis the stability measure
-obtained with FastSRM (identifiable since the shared components covariance is diagonal).
+We plot the results as a scatter plot in
+Figure~\ref{exp:identifiability}, where, for each repetition, the
+x-axis indicates the stability measure obtained with Brainiak's
+implementation and the y-axis the stability measure obtained with
+FastSRM (identifiable since the shared components covariance is
+diagonal).
 %
@@
 We estimate the shared responses by averaging the individual components of each
-subjects but one.
+subject but one.
 %
@@
 \section{Conclusion}
-As studies using naturalistic stimuli will tend to become more common and large
+As studies using naturalistic stimuli will tend to become more common
 within and across subjects, we need scalable models in terms of computation time
@@
 
-%
 FastSRM is an efficient implementation of SRM
@@
 
-%
 FastSRM inherits from all the weaknesses of SRM including the fact that
-mixing matrices are assumed to be orthogonal, which is debattable in practice. It
-remains to be seen whether dimension reduction and learning of the shared
+mixing matrices are assumed to be orthogonal, which is debattable in practice.
+%
+It is open question whether dimension reduction and learning of the shared
 components could be done jointly and efficiently without assuming orthogonal
@@
 Our approach based on optimal spatial decomposition shows that SRM-like methods do not need the
-full data to provide accurate estimates of shared components. We believe such
+full data to provide accurate estimates of shared components. We believe that such
 insights may guide the design of future methods.
-%
-
-Lastly, FastSRM is useful when the number of features is much larger than the
+
+Last, FastSRM is useful when the number of features is much larger than the
 number of samples. This setting is common in fMRI but less common in MEG/EEG where
@@
 
+\section{Detailed derivation of the Probabilistic SRM algorithm}
+\label{details:probsrm}
+Here we describe in detail the log-likeihood underlying the
+Probabilistic SRM and derive the update rules.
+%
+Denoting $\VV[\sbb | \xb] = (\sum_i \frac1{\sigma_i^2} I +
+\Sigma_s^{-1})^{-1}$ and $\EE[\sbb | \xb] = \VV[\sbb | \xb] \sum_i \frac1{\sigma_i^2}
+A_i^{\top}\xb_i$, we have
+
+\begin{align}
+  p(\xb, \sbb) &= \prod_i \frac{\exp \left(-\frac{\|\xb_i - A_i \sbb \|^2}{2 \sigma_i^2} \right) }{(2 \pi \sigma_i^{2v})^{\frac12}} \frac{\exp(-\frac12 \dotp{ \sbb }{ \Sigma_s^{-1} \sbb } )}{(2 \pi | \Sigma_s|)^{\frac12}} \\
+               &= c_1 \exp (-\frac12 \left( \sum_i \frac1{\sigma_i^2}\|\xb_i\|^2 - 2  \dotp{ \sum_i \frac1{\sigma_i^2} A_i^{\top}\xb_i}{ \sbb } \right.
+\\& \left.+ \sum_i \frac1{\sigma_i^2} \| \sbb \|^2 + \dotp{ \sbb}{ \Sigma_s^{-1} \sbb }  \right)) \\
+               &= c_2(\xb) \exp(-\frac12 \left( \dotp{  \sbb - \EE[\sbb }{ \xb], \VV[\sbb | \xb]^{-1} ( \sbb - \EE[\sbb | \xb])  } \right)) \label{probsrmcompleted}
+\end{align}
+where
+\begin{align}
+  c_1 = \frac1{(2 \pi \sigma_i^{2v})^{\frac12}}\frac1{(2 \pi |
+  \Sigma_s|)^{\frac12}}
+\end{align}
+and
+\begin{align}
+  c_2(\xb) = c_1 \exp \left( -\frac12 \left( \sum_i \frac1{\sigma_i^2}\|\xb_i\|^2 - \dotp{  \EE[\sbb }{ \xb], \VV[\sbb | \xb]^{-1} \EE[\sbb | \xb] } \right) \right)
+\end{align}
+are independent from $\sbb$.
+%
+Therefore,
+\begin{align}
+  \sbb| \xb \sim \Ncal(\EE[\sbb | \xb], \VV[\sbb, \xb])
+\end{align}
+
+The negative expected completed log-likelihood is given by
+\begin{align}
+	\loss = \sum_i \frac12 v \log(\sigma_i^2) + \frac1{2 \sigma_i^2} \EE[\| \xb_i - A_i \sbb \|^2]
+\end{align}
+updates are therefore given by:
+\begin{align}
+  &\sigma_i^2 \leftarrow \frac1{v} (\EE[\| \xb_i - A_i \EE[\sbb|\xb]\|^2] + \| \diag(\VV[\sbb | \xb]) \|) \\
+  % no square here, right ?
+  &A_i \leftarrow \Pcal(\EE[\xb_i \EE[\sbb|\xb]^{\top}]) \label{eq:psrm:Aiupdate} \\
+  & \Sigma_s \leftarrow \VV[\sbb | \xb] + \EE[\EE[\sbb | \xb] \EE[\sbb | \xb]^{\top}]
+\end{align}
+
+It is useful to access the log-likelihood to check the implementation of the
+algorithm and monitor the convergence.
+%
+From equation~\eqref{probsrmcompleted},
+the likelihood is given by:
+\begin{align}
+  p(\xb) &= c_2(\xb) \int_{\sbb} \exp(-\frac12 \left( \dotp{  \sbb - \EE[\sbb }{ \xb], \VV[\sbb | \xb]^{-1} ( \sbb - \EE[\sbb | \xb])  } \right)) d\sbb \\
+         &= c_2(\xb) (2 \pi |\VV[\sbb | \xb]|)^{\frac12}
+\end{align}
+replacing $c_2(\xb)$ by its expression and taking the log, the expected negative
+log-likelihood is (up to constants) given by:
+\begin{align}
+  \EE[-\log(&p(\xb))] = \sum_i \frac{v}{2} \log(\sigma_i^2) + \frac12 \log(|\Sigma_s|) - \frac12 \log(|\VV[\sbb | \xb]|) \nonumber \\ &+ \sum_i
+  \frac12 \frac1{\sigma_i^2} \EE[\|\xb_i\|^2] - \frac12 \EE[\dotp{  \EE[\sbb }{ \xb], \VV[\sbb | \xb]^{-1} \EE[\sbb | \xb] }]
+\end{align}
+
 \section{Proofs of Proposition~\ref{prop:optimaldetsrm}}

### 2022-01-04 paper/main.tex
correct small typos

@@
 %
-Although SHERLOCK originally data contain 1 run only, we split them into 4 blocks
+Although SHERLOCK data originally contain 1 run only, we split them into 4 blocks
 of 395 time frames and one block of 396 time frames for the needs of our
@@
 %
-It is open question whether dimension reduction and learning of the shared
+It is an open question whether dimension reduction and learning of the shared
 components could be done jointly and efficiently without assuming orthogonal

### 2022-01-04 paper/main.tex
correct abstract

@@
 \begin{abstract}
-  The shared response model (SRM) provides a simple but effective framework to analyse
+  The shared response model (SRM) provides a simple but effective framework to analyze
   fMRI data of subjects exposed to naturalistic stimuli.

### 2022-01-04 paper/main.tex
typos

@@
 %
-We then introduce FastSRM, that relies on a dimension reduction step that we
+We then introduce FastSRM, which relies on a dimension reduction step that we
 prove to yield the same solution as the original algorithm.
@@
 
-The experiments performed in the paper are \emph{fully} reproducible: our
-code available at \url{https://github.com/hugorichard/FastSRM} allows to download the
+The experiments performed in this article are \emph{fully} reproducible: our
+code available at \url{https://github.com/hugorichard/FastSRM} allows you to download the
 data, run the experiments and plot the figures.

### 2022-01-07 fastsrm/identifiable_srm.py
sensible defaults

@@ -139,7 +139,7 @@ Fast shared response model for fMRI data (https://arxiv.org/pdf/1909.12537.pdf)
         self,
         method="prob",
         n_components=20,
-        n_iter=1,
+        n_iter=1000,
         temp_dir=None,
         n_jobs=1,
         verbose="warn",

### 2022-01-07 tests/test_identifiable_srm.py
correct mistake in one test

@@ -352,6 +352,7 @@ def test_use_pca(method):
         srm2 = IdentifiableFastSRM(
             n_components=5,
             use_pca=True,
+            method=method
         )
         B = srm2.fit_transform(X_train)
         np.testing.assert_array_almost_equal(A, B, 4)

### 2022-06-14 paper/main.tex
update stuff

@@
 Lastly, another issue of SRM is its non-identifiability (see the proof
-in~\cite{richard2020modeling} Appendix D) which makes it challenging to
+in~\cite{richard2020modeling} Appendix D also reported in Appendix~\ref{app:unidentifiability} for completness) which makes it challenging to
 interpret the extracted shared components as they are defined up to a rotation.
@@
 FastSRM gains identifiability by imposing that the covariance of the shared
-components is diagonal (see ~\cite{richard2020modeling} Appendix D for a proof).
+components is diagonal (see ~\cite{richard2020modeling} Appendix D or Appendix~\ref{app:unidentifiability} of this paper for a proof).
 %
@@
 As already highlighted in introduction, this causes the model to be
-unidentifiable (see~\cite{richard2020modeling} Appendix D for a proof). Enforcing a diagonal $\Sigma_s$ ensures identifiability, provided that the diagonal values are different.
+unidentifiable (see~\cite{richard2020modeling} Appendix D or Appendix~\ref{app:unidentifiability} of this paper for a proof).
+Enforcing a diagonal $\Sigma_s$ ensures identifiability, provided that the diagonal values are different.
 %
@@
 to~\cite{chen2015reduced} and~\cite{anderson2016enabling} for the original formulation of Probabilistic SRM
-without the diagonal constraint)
+without the diagonal constraint).
+The following paragraph gives update rules for $\lambda_i$ and $A_i$ and $\Sigma_s$.
+We highlight that the derivations used to obtain these updates are not new.
+They are obtained following the strategy introduced in~\cite{anderson2016enabling} and~\cite{chen2015reduced}.
 %
@@
   \centering
-  \includegraphics[scale=0.34]{figures/srm/conceptual_figure2.pdf}
+  \includegraphics[scale=0.34]{figures/srm/conceptual_figure2_revised.pdf}
   \caption{\textbf{FastSRM algorithm} In step 1, data $\xb_i$ are projected onto a
@@
   Let $(A_i)_i, \sigma_i, \Sigma_s$ be the solution obtained by probabilistic SRM on data
-  $\xb_i$ and $(A_i')_i, \sigma_i', \Sigma_s'$ the solution obtained by ProbSRM on
-  data $\zb_i = U_{\xb_i}^{\top} \xb_i$ but where updates:
-  \begin{align}
-    \sigma_i^2 \leftarrow \frac1{t} (\EE[\| \zb_i - A_i \EE[\sbb|\zb]\|^2] + \| \diag(\VV[\sbb | \zb]) \|^2)
+  $\xb_i \in \RR^v$ and $(A_i')_i, \sigma_i', \Sigma_s'$ the solution obtained by ProbSRM on
+  data $\zb_i = U_{\xb_i}^{\top} \xb_i$, $\zb_i \in \RR^n$ but where updates:
+  \begin{align}
+    \sigma_i^2 \leftarrow \frac1{n} (\EE[\| \zb_i - A_i \EE[\sbb|\zb]\|^2] + \| \diag(\VV[\sbb | \zb]) \|^2)
   \end{align}
@@
   %
+  where $t$ is the number of samples and $v$ is the number of voxels.
+
   Then $A_i = U_{\xb_i}A_i'$, $\sigma_i =
@@
 %
-In this work, we do not discuss the merit of FastSRM with respect to
-other methods. We simply observe that SRM is widely used and aim to provide
+SRM has been introduced after hyper-alignment in~\cite{chen2015reduced} and has been shown to outperform hyperalignment in many different settings.
+A recent work has come to a similar conclusion~\cite{bazeille2021empirical}.
+In this article, we do not discuss further the merit of FastSRM with respect to other methods.
+We simply observe that SRM is widely used and aim to provide
 a faster, more memory efficient and identifiable implementation.  In
@@
 with provably no loss of performance.
-Note that the new memory requirements do not depend on the number of subjects.
-This is really interesting as it would make it possible to apply FastSRM on
-datasets with a very large number of subjects that do not hold in memory.
 
@@
 mixing matrices are assumed to be orthogonal, which is debattable in practice.
+Another desirable property of such decompositions is the spatial smoothness of the components.
+While it is often observed that SRM components are smooth, nothing imposes such a constraint in the model.
+We highlight that the optimal spatial decomposition is only optimal for the model we consider.
+Whether this representation is interpretable or not is irrelevant as it is only an intermediate step produced used during training.
+
 %
@@
 Probabilistic SRM and derive the update rules.
+We highlight that the derivations used to obtain these updates are not new. They are obtained following the strategy introduced in~\cite{anderson2016enabling} and~\cite{chen2015reduced}.
+
 %
@@
 \end{align}
+
+\section{Identifiability results on Probabilistic SRM available in~\cite{richard2020modeling} appendix D}
+\label{app:unidentifiability}
+\begin{proposition}
+Probabilistic SRM with positive definite $\Sigma_s$ is not identifiable
+\end{proposition}
+\begin{proof}
+Let us assume the data $\xb_i \enspace i=1, \dots, m$ follow the Probabilistic SRM model with parameters $\Sigma_s, A_i, \sigma_i^2 \enspace i=1, \dots, m$ and $\Sigma_s$ definite positive.
+
+Let us consider an orthogonal matrix $O \in \Ocal_k$.
+We call $A'_i = A_i O$ and $\Sigma'_s = O^{\top} \Sigma_s O$.
+$\Sigma'_s$ is trivially symmetric positive definite.
+
+Then the data also follows the SRM model with different parameters $\Sigma'_s, A'_i, \sigma_i^2 \enspace i=1, \dots, m$.
+\end{proof}
+
+\begin{proposition}
+We consider the Probabillistic SRM model where $\Sigma_s$ is a positive \emph{diagonal} matrix. We further assume that the values in $\Sigma$ are all distinct and ranked in ascending order.
+Then, the Probabilistic SRM model is identifiable up to sign indeterminacies on the columns of
+$\begin{bmatrix}
+A^1 \\
+\vdots \\
+A^m
+\end{bmatrix}
+$.
+\end{proposition}
+\begin{proof}
+The Probabilistic SRM can be written
+\begin{align*}
+    &\xb_i \sim \Ncal(0, A_i \Sigma_s {A_i}^{\top} + \sigma_i^2 I_v) \enspace \text{with}\enspace  {A_i}^{\top}A_i = I_p
+\end{align*}
+where $\Sigma_s$ is a positive diagonal matrix with distincts values ranked in ascending order.
+
+Let us assume the data $\xb^i \enspace i=1, \dots, m$ follow the Probabilistic SRM model with parameters $\Sigma_s, A_i, {\sigma_i}^2 \enspace i=1, \dots, m$.
+Let us further assume that the data $\xb^i \enspace i=1, \dots, m$ follow the Probabilistic SRM model with an other set of parameters $\Sigma'_s, A'_i, {\sigma'_i}^2 \enspace i=1, \dots, m$.
+
+Since the model is Gaussian, we look at the covariances.
+We have for $i \neq j$
+\begin{align*}
+    \bbE[\xb_i\left(\xb_j\right)^{\top}] = A_i\Sigma_s {A_j}^{\top} = A'_i \Sigma'_s{A'_j}^{\top} \enspace,
+\end{align*}
+The singular value decomposition is unique up to sign flips and permutation.
+Since eigenvalues are positive and ranked the only indeterminacies left are on the eigenvectors.
+For each eigenvalue a sign flip can occur simultaneously on the corresponding left and right eigenvector.
+
+Therefore we have $\Sigma'_s = \Sigma_s$, $A_i = A'_i D_{ij}$ and $A_j = A'_j D_{ij}$ where $D_{ij} \in \bbR^{k, k}$ is a diagonal matrix with values in $\{-1, 1\}$.
+This analysis holds for every $j \neq i$ and therefore $D_{ij} = D$ is the same for all subjects.
+
+We also have for all $i$
+\begin{align*}
+    \bbE[\xb_i \left(\xb_i\right)^{\top}] = A_i \Sigma_s {A_i}^{\top} + \sigma_i^2 I_v =  A'_i \Sigma'_s {A'_i}^{\top}  + {\sigma}_i^2 I_v\\
+\end{align*}
+We therefore conclude ${\sigma}_i^2 = \sigma_i^2, i=1 \dots m$.
+
+Note that if the diagonal subject specific noise covariance $\sigma_i^2 I_v$ is replaced by any positive definite matrix, the model still enjoys identifiability.
+\end{proof}
 
@@
   \begin{align}
-    \sigma_i^2 \leftarrow \frac1{t} (\EE[\| \zb_i - \tilde{A}_i \EE[\sbb|\zb]\|^2] + \| \diag(\VV[\sbb | \zb]) \|^2)
+    \sigma_i^2 \leftarrow \frac1{n} (\EE[\| \zb_i - \tilde{A}_i \EE[\sbb|\zb]\|^2] + \| \diag(\VV[\sbb | \zb]) \|^2)
   \end{align}

### 2022-07-28 paper/main.tex
edited the intro

@@
 prove to yield the same solution as the original algorithm.
-% be optimal in the sense that working with reduced data does not induce any change in the
-% algorithm trajectory.
 %
@@
 %
-Yet, in contrast to standard designs, these so-called
-naturalistic stimuli do not come with a design matrix that quantifies the
-features of interest in the stimuli.
-%
-Without a design matrix, models such as the general linear
-model~\cite{poline2012general} cannot be used.
+If one wants to use data acquired under naturalistic conditions to map
+functional responses to the brain, i.e. in an encoding setting
+\cite{naselaris2011}, one has to deal with the fact that no design matrix is
+specified for naturalistic stimuli.
+%
+Such a design matrix encodes the temporal events that affect brain
+signal during an experiment; these events typically reflect the occurrence
+of some features of interest in the stimuli.
 %
@@
 design matrix of naturalistic stimuli, it is a hard task.
-
-Another approach is to learn jointly the design matrix and the spatial maps in
+%
+Without a design matrix, models such as the general linear
+model~\cite{poline2012general} cannot be used.
+
+Another approach is to learn jointly a set of reference time courses  and the spatial maps in
 an unsupervised way.
+%
+These reference time courses formally replace the design matrix in the
+spatio-temporal decomposition of the data, but cannot be considered as
+a formal design matrix.
 %
@@
 When the dataset contains multiple subjects undergoing the same
-protocol, a natural assumption is to assume that the design matrix is
+protocol, a natural assumption is to assume that the temporal reference is
 shared across subjects.

### 2022-08-03 paper/main.tex
update intro

@@
   %
-  where $t$ is the number of samples and $v$ is the number of voxels.
+  where $n$ is the number of samples and $v$ is the number of voxels.
 
@@
 SRM has been introduced after hyper-alignment in~\cite{chen2015reduced} and has been shown to outperform hyperalignment in many different settings.
-A recent work has come to a similar conclusion~\cite{bazeille2021empirical}.
+Specifically, SRM is natively a group model, which hyperalignment is not, it is more computationally efficient, with the avoidance of searchlight loops, and sometimes obtains higher accuracy than Procrustes hyperalignment~\cite{bazeille2021empirical}.
 In this article, we do not discuss further the merit of FastSRM with respect to other methods.
-We simply observe that SRM is widely used and aim to provide
-a faster, more memory efficient and identifiable implementation.  In
-our view, FastSRM should be used as a plugin replacement for SRM.
+We simply observe that SRM is widely used and aim to provide a faster, more memory efficient and identifiable implementation.
+In our view, FastSRM should be used as a plugin replacement for SRM.
 
@@
 % %
-In general probabilistic methods give much better results than their deterministic
-counterpart.
-%
-This shows the superiority of likelihood based methods.
+We also see that ProbSRM gives much better results than DetSRM.
 %
@@
 Assuming that the data follow the model, identifiability ensures that the
-problem is well-posed since the solutions are well-characterized. Note however that SRM
+problem is well-posed since the solutions are well-characterized.
+More precisely, if the data are generated according to an identifiable model, in the limit of infinite samples, a perfect optimization algorithm is guaranteed to recover the parameters used to generate the data.
+In contrast, in an unidentifiable model, there are more than one set of parameters that can generate the same data making the search for the good parameters ill-defined.
+In practice, unidentifiable models have a more complex optimization landscape with many more local minima than identifiable models (any local minima in the identifiable model produces as many local minimas as they are indeterminacies in the corresponding unidentifiable model).
+This is why the unidentifiable models are therefore expected to be more sensitive to initialization and less robust to small changes in the data than their identifiable counterparts.
+Note however that SRM
 remains non-convex so there is no guarantee that the optimal solution is found.
+
 %
@@
 %
-The components are then matched using the Hungarian algorithm and the stability
+The components are then matched using the Hungarian algorithm~\cite{kuhn1955hungarian} and the stability
 measure is obtained from the average correlation of the matched components.
@@
 While it is often observed that SRM components are smooth, nothing imposes such a constraint in the model.
-We highlight that the optimal spatial decomposition is only optimal for the model we consider.
-Whether this representation is interpretable or not is irrelevant as it is only an intermediate step produced used during training.
-
+We highlight that the optimal spatial decomposition is only optimal with respect to the SRM model.
+More precisely, it gives a reduced representation such that applying the SRM model on the reduced representation or on the full data gives the same result.
+Whether this representation is interpretable or not is irrelevant to us, since it is only a means to accelerate the SRM algorithm.
+However, as this representation is not unique, it may be interesting to find the “most interpretable” optimal spatial decomposition.
+Note that this would not change the end result of the SRM algorithm but could be of interest for interpretation purposes.
 %

### 2022-08-04 paper/main.tex
pass by BT

@@
   of subjects or runs is large,
-  % \bt{True, but we the experiments don't make it so explicit ?}
   fitting the model requires a large amount of memory
@@
 %
-We then introduce FastSRM, which relies on a dimension reduction step that we
-prove to yield the same solution as the original algorithm.
+We then introduce FastSRM, that relies on a dimension reduction step and yields the same solution as the original algorithm.
 %
@@
 spatio-temporal decomposition of the data, but cannot be considered as
-a formal design matrix.
-%
-In the case where the dataset contains only one subject,
+a design matrix, as it is data-driven.
+%
+In the case where the dataset only contains  one subject,
 independent component analysis~\cite{jutten1991blind} (ICA) is the method of choice.
@@
 Lastly, another issue of SRM is its non-identifiability (see the proof
-in~\cite{richard2020modeling} Appendix D also reported in Appendix~\ref{app:unidentifiability} for completness) which makes it challenging to
+in~\cite{richard2020modeling} Appendix D also reported in Appendix~\ref{app:unidentifiability} for completeness) which makes it challenging to
 interpret the extracted shared components as they are defined up to a rotation.
@@
   \centering
-  \includegraphics[scale=0.3]{figures/srm/conceptual_figure33.png}
+  \includegraphics[scale=0.4]{figures/srm/conceptual_figure33.png}
   \caption{\textbf{Shared response model}: The raw fMRI data are modeled as a
@@
 Indeed, after the updates with respect to each mixing matrix have been
-carried out, the gradient with respect to $(W_i)_{i=1}^m$ is zero and therefore
+carried out, the gradient with respect to $(W_i)_{i=1}^m$ \textbf{XXX $(W_i)_{i=1}^m$ not introduced so far} is zero and therefore
 to compute the $\ell_{\infty}$ norm of the gradient we only need the gradient
@@
 The number of components in large spatial decompositions is about $r=1000$ and in full brain data,
-the number of voxels is about $300~000$ so that $\frac{v}{r}$ is typically about
-$300$.
+the number of voxels is typically in the  $50~000 - 400~000$ range, so that $\frac{v}{r}$ is typically about
+$50$ to $400$.
 %
@@
 
-We empirically see that the optimal approach is equivalent to using no spatial decompositions in
+We empirically see that the optimal approach is equivalent to using no spatial decomposition in
 terms of performance.
@@
 Our work is fully reproducible and our code is tested and well documented. All
-material is available in our github repository \url{https://github.com/hugorichard/FastSRM}.
+material is available in the \url{https://github.com/hugorichard/FastSRM} Github repository.
 
@@
 
-While brainiak and nilearn are useful to reproduce some of our experiments, they
+While Brainiak and Nilearn are useful to reproduce some of our experiments, they
 are not strong dependencies and FastSRM only needs Numpy ($\geq 1.12$), Scipy
@@
 While it is often observed that SRM components are smooth, nothing imposes such a constraint in the model.
+%
 We highlight that the optimal spatial decomposition is only optimal with respect to the SRM model.
+%
 More precisely, it gives a reduced representation such that applying the SRM model on the reduced representation or on the full data gives the same result.
-Whether this representation is interpretable or not is irrelevant to us, since it is only a means to accelerate the SRM algorithm.
+%
+Whether this representation is interpretable or not is irrelevant here, since we only focus on accelerating the SRM algorithm.
+%
 However, as this representation is not unique, it may be interesting to find the “most interpretable” optimal spatial decomposition.

### 2022-08-06 paper/main.tex
replace W_i by A_i

@@
 Indeed, after the updates with respect to each mixing matrix have been
-carried out, the gradient with respect to $(W_i)_{i=1}^m$ \textbf{XXX $(W_i)_{i=1}^m$ not introduced so far} is zero and therefore
+carried out, the gradient with respect to $(A_i)_{i=1}^m$ is zero and therefore
 to compute the $\ell_{\infty}$ norm of the gradient we only need the gradient
